---
title: "WRF Output SpBayes"
author: "Renata Poulton Kamakura"
date: "2025-02-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(spBayes) #spatial random effects

#for data reformatting and spBayes setup
library(sf)
library(sp)
library(dplyr)
#library(parallel) #make it run faster

library(RANN) #for knots

# Calculate the number of cores
no_cores <- 3

```

##Read in data

Creating the LU dictionary to simplify the land use categories a bit for the run, we are not as worried about subtleties of LU impacts

For calculating wind direction: https://stackoverflow.com/questions/21484558/how-to-calculate-wind-direction-from-u-and-v-wind-components-in-r
https://confluence.ecmwf.int/pages/viewpage.action?pageId=133262398 

```{r read data}
#Get the hourly averages for each year
#Likely will need to update the files here once re-run with all the data and spin-up
HourlyMeans16 <- read.csv("../../03ProcessedData/HourlyMeans_2016_051325.csv")
HourlyMeans17 <- read.csv("../../03ProcessedData/HourlyMeans_2017_051325.csv")
HourlyMeans17_msst <- read.csv("../../03ProcessedData/HourlyMeans_2017_w16SST_051325.csv")

#More for reference
LU_names = c('Evergreen Needleleaf Forest', 'Evergreen Broadleaf Forest', 
            'Deciduous Needleleaf Forest', 'Deciduous Broadleaf Forest',
            'Mixed Forests', 'Closed Shrublands',  'Open Shrublands', 
            'Woody Savannas', 'Savannas', 'Grasslands', 
            'Permanent wetlands', 'Croplands', 'Urban and Built-Up', 
            'cropland/natural vegetation mosaic', 
            'Snow and Ice', 'Barren or Sparsely Vegetated', 'Water')

NewLUs <- c('Forest', 'Forest', 'Forest', 'Forest', 'Forest',
            'Shrub/Savanna', 'Shrub/Savanna', 'Shrub/Savanna', 'Shrub/Savanna',
            "Shrub/Savanna", "Wetlands", "Crop", "Urban", "Crop", "Snow", 
            "Barren", "Water")
NewLU_nums <- c(1,1,1,1,1,2,2,2,2,2,3,4,5,4,6,7,8)

LU_dict <- as.data.frame(matrix(0, nrow=17, ncol=4))
names(LU_dict) <- c("Original_Name", "Original_Num", "New_Name", "New_Num")

LU_dict$Original_Name <- LU_names
LU_dict$Original_Num <- 1:17
LU_dict$New_Name <- NewLUs
LU_dict$New_Num <- NewLU_nums

```


##Start in on regressions

So we would need to account for a few things
* spatial random effects
* fixed effects
  - land use
  - distance from ocean
  - hour
  
The relationship should change throughout the day, you see that the slopes of some of these relationships shift over time...
* hour x land use (really key)
* hour x distance2ocean? 

Alternatively, could run a few separate models at different times of day so you avoid some of the interactive effects
* early afternoon (12-14)
* peak temp (15-17)
* evening (18-21)

spBayes information
* https://faculty.ucr.edu/~jflegal/203/spBayes_tutorial.pdf
* https://cran.r-project.org/web/packages/spBayes/spBayes.pdf

If you do spDynLm I think you end up with different regressions for each hour
* so you end up with temps at each hour as your y variables
* then you coordinates are your spatial random effects
* your fixed effects are then land use and distance2ocean


```{r create knots}

#this should work for both datasets
#nobs <- length(HourlyMeans17$GridCell) #number of observations
#coords <- HourlyMeans17[, c("GridCell", "lat", "long")] #points of the observations

nobs <- length(HourlyMeans17_msst$GridCell)
coords <- HourlyMeans17_msst[, c("GridCell", "lat", "long")]
names(coords) <- c("GridCell", "Y", "X") #yes I know X should be longitude, but I somehow flipped it earlier
coords <- unique(coords)

#change your coordinates to UTM so the distances are consistent
coords_sf <- st_as_sf(coords, coords = c("X", "Y"), crs = 4326)
coords_UTM <- st_transform(coords_sf, CRS("+proj=utm +zone=15 ellps=WGS84"))


##add knots to make this run faster - dimension reduction
#https://rpubs.com/jimclark/751504
nk   <- 12
boundingbox <- st_bbox(coords_UTM)
klon <- seq(boundingbox[1],boundingbox[3],length=nk)[-c(1,nk)]
klat <- seq(boundingbox[2],boundingbox[4],length=nk)[-c(1,nk)]
knots <- as.matrix( expand.grid(klon,klat) )
kdf <- as.data.frame(knots)
names(kdf) <- c("X", "Y")
knots_sf <- st_as_sf(kdf, coords=c("X", "Y"), crs = CRS("+proj=utm +zone=15 ellps=WGS84"))

knots_test <- knots_sf %>% 
  mutate(within_2000 = lengths(st_is_within_distance(x = .,
                                                    y = coords_UTM,
                                                    dist = 2000)))

ntot <- 50 #number of knots to keep
knots <- st_coordinates(knots_test$geometry[order(knots_test$within_2000, decreasing = TRUE)][1:ntot])

#check the knots
plot(boundingbox, xlim=c(boundingbox[1], boundingbox[3]), ylim=c(boundingbox[2], boundingbox[4]))
points(st_coordinates(coords_UTM),cex=.2)
points(knots[,1],knots[,2],col='black', pch=15)

```

Now that we have some spatial dimension reduction, re-organize the data into the format we need

```{r reformat data}
#reduce your landuse categories
HourlyMeans16$LandUse <- LU_dict$New_Name[HourlyMeans16$LandUse]
HourlyMeans17$LandUse <- LU_dict$New_Name[HourlyMeans17$LandUse]
HourlyMeans17_msst$LandUse <- LU_dict$New_Name[HourlyMeans17_msst$LandUse]

#Should set your baseline land use - maybe to forest?
HourlyMeans16$LandUse <- as.factor(HourlyMeans16$LandUse)
HourlyMeans16$LandUse <- relevel(HourlyMeans16$LandUse, ref="Crop")
HourlyMeans17$LandUse <- as.factor(HourlyMeans17$LandUse)
HourlyMeans17$LandUse <- relevel(HourlyMeans17$LandUse, ref="Crop")
HourlyMeans17_msst$LandUse <- as.factor(HourlyMeans17_msst$LandUse)
HourlyMeans17_msst$LandUse <- relevel(HourlyMeans17_msst$LandUse, ref="Crop")

HoursOfInterest <- 12:21 #partly from EDA, basically going from approaching peak temps and down into the evening and the start of cool-down

n16 <- length(unique(HourlyMeans16$GridCell)) #observations per hour
n17 <- length(unique(HourlyMeans17$GridCell)) #observations per hour
n17c <- length(unique(HourlyMeans17_msst$GridCell)) #observations per hour

################################2016: Non-MHW
#Re-organize 2016 - Temperature
#columns are one for each hour, one for distance2ocean, and one for land use
ReOrg_Temp16 <- as.data.frame(matrix(0, ncol=(length(HoursOfInterest) + 2), nrow = n16))

HourlyMeans16 <- HourlyMeans16[order(HourlyMeans16$GridCell),] #make sure the order is consistent so you know which points are which

#Get your hourly temps - response
ReOrg_Temp16[,1] <- HourlyMeans16$X2mTemp[HourlyMeans16$hour == 12]
ReOrg_Temp16[,2] <- HourlyMeans16$X2mTemp[HourlyMeans16$hour == 13]
ReOrg_Temp16[,3] <- HourlyMeans16$X2mTemp[HourlyMeans16$hour == 14]
ReOrg_Temp16[,4] <- HourlyMeans16$X2mTemp[HourlyMeans16$hour == 15]
ReOrg_Temp16[,5] <- HourlyMeans16$X2mTemp[HourlyMeans16$hour == 16]
ReOrg_Temp16[,6] <- HourlyMeans16$X2mTemp[HourlyMeans16$hour == 17]
ReOrg_Temp16[,7] <- HourlyMeans16$X2mTemp[HourlyMeans16$hour == 18]
ReOrg_Temp16[,8] <- HourlyMeans16$X2mTemp[HourlyMeans16$hour == 19]
ReOrg_Temp16[,9] <- HourlyMeans16$X2mTemp[HourlyMeans16$hour == 20]
ReOrg_Temp16[,10] <- HourlyMeans16$X2mTemp[HourlyMeans16$hour == 21]

#Get your fixed effects
ReOrg_Temp16[,11] <- HourlyMeans16$Dist2Ocean[HourlyMeans16$hour == 12]
ReOrg_Temp16[,12] <- HourlyMeans16$LandUse[HourlyMeans16$hour == 12]
names(ReOrg_Temp16) <- c("y.1", "y.2", "y.3", "y.4", "y.5", "y.6", "y.7", "y.8", "y.9", "y.10", "Dist2Ocean", "LandUse")


#re-organize 2016 humidity
ReOrg_RH16 <- as.data.frame(matrix(0, ncol=(length(HoursOfInterest) + 2), nrow =n16))

#Get your hourly temps - response
ReOrg_RH16[,1] <- HourlyMeans16$RelHumid[HourlyMeans16$hour == 12]
ReOrg_RH16[,2] <- HourlyMeans16$RelHumid[HourlyMeans16$hour == 13]
ReOrg_RH16[,3] <- HourlyMeans16$RelHumid[HourlyMeans16$hour == 14]
ReOrg_RH16[,4] <- HourlyMeans16$RelHumid[HourlyMeans16$hour == 15]
ReOrg_RH16[,5] <- HourlyMeans16$RelHumid[HourlyMeans16$hour == 16]
ReOrg_RH16[,6] <- HourlyMeans16$RelHumid[HourlyMeans16$hour == 17]
ReOrg_RH16[,7] <- HourlyMeans16$RelHumid[HourlyMeans16$hour == 18]
ReOrg_RH16[,8] <- HourlyMeans16$RelHumid[HourlyMeans16$hour == 19]
ReOrg_RH16[,9] <- HourlyMeans16$RelHumid[HourlyMeans16$hour == 20]
ReOrg_RH16[,10] <- HourlyMeans16$RelHumid[HourlyMeans16$hour == 21]

#Get your fixed effects
ReOrg_RH16[,11] <- HourlyMeans16$Dist2Ocean[HourlyMeans16$hour == 12]
ReOrg_RH16[,12] <- HourlyMeans16$LandUse[HourlyMeans16$hour == 12]
names(ReOrg_RH16) <- c("y.1", "y.2", "y.3", "y.4", "y.5", "y.6", "y.7", "y.8", "y.9", "y.10", "Dist2Ocean", "LandUse")

############################# 2017: MHW
#Re-organize 2017 - Temperature
#columns are one for each hour, one for distance2ocean, and one for land use
ReOrg_Temp17 <- as.data.frame(matrix(0, ncol=(length(HoursOfInterest) + 2), nrow =n17))

HourlyMeans17 <- HourlyMeans17[order(HourlyMeans17$GridCell),] #make sure the order is consistent so you know which points are which

#Get your hourly temps - response
ReOrg_Temp17[,1] <- HourlyMeans17$X2mTemp[HourlyMeans17$hour == 12]
ReOrg_Temp17[,2] <- HourlyMeans17$X2mTemp[HourlyMeans17$hour == 13]
ReOrg_Temp17[,3] <- HourlyMeans17$X2mTemp[HourlyMeans17$hour == 14]
ReOrg_Temp17[,4] <- HourlyMeans17$X2mTemp[HourlyMeans17$hour == 15]
ReOrg_Temp17[,5] <- HourlyMeans17$X2mTemp[HourlyMeans17$hour == 16]
ReOrg_Temp17[,6] <- HourlyMeans17$X2mTemp[HourlyMeans17$hour == 17]
ReOrg_Temp17[,7] <- HourlyMeans17$X2mTemp[HourlyMeans17$hour == 18]
ReOrg_Temp17[,8] <- HourlyMeans17$X2mTemp[HourlyMeans17$hour == 19]
ReOrg_Temp17[,9] <- HourlyMeans17$X2mTemp[HourlyMeans17$hour == 20]
ReOrg_Temp17[,10] <- HourlyMeans17$X2mTemp[HourlyMeans17$hour == 21]

#Get your fixed effects
ReOrg_Temp17[,11] <- HourlyMeans17$Dist2Ocean[HourlyMeans17$hour == 12]
ReOrg_Temp17[,12] <- HourlyMeans17$LandUse[HourlyMeans17$hour == 12]
names(ReOrg_Temp17) <- c("y.1", "y.2", "y.3", "y.4", "y.5", "y.6", "y.7", "y.8", "y.9", "y.10", "Dist2Ocean", "LandUse")


#re-organize 2017 humidity
ReOrg_RH17 <- as.data.frame(matrix(0, ncol=(length(HoursOfInterest) + 2), nrow =n17))

#Get your hourly temps - response
ReOrg_RH17[,1] <- HourlyMeans17$RelHumid[HourlyMeans17$hour == 12]
ReOrg_RH17[,2] <- HourlyMeans17$RelHumid[HourlyMeans17$hour == 13]
ReOrg_RH17[,3] <- HourlyMeans17$RelHumid[HourlyMeans17$hour == 14]
ReOrg_RH17[,4] <- HourlyMeans17$RelHumid[HourlyMeans17$hour == 15]
ReOrg_RH17[,5] <- HourlyMeans17$RelHumid[HourlyMeans17$hour == 16]
ReOrg_RH17[,6] <- HourlyMeans17$RelHumid[HourlyMeans17$hour == 17]
ReOrg_RH17[,7] <- HourlyMeans17$RelHumid[HourlyMeans17$hour == 18]
ReOrg_RH17[,8] <- HourlyMeans17$RelHumid[HourlyMeans17$hour == 19]
ReOrg_RH17[,9] <- HourlyMeans17$RelHumid[HourlyMeans17$hour == 20]
ReOrg_RH17[,10] <- HourlyMeans17$RelHumid[HourlyMeans17$hour == 21]

#Get your fixed effects
ReOrg_RH17[,11] <- HourlyMeans17$Dist2Ocean[HourlyMeans17$hour == 12]
ReOrg_RH17[,12] <- HourlyMeans17$LandUse[HourlyMeans17$hour == 12]
names(ReOrg_RH17) <- c("y.1", "y.2", "y.3", "y.4", "y.5", "y.6", "y.7", "y.8", "y.9", "y.10", "Dist2Ocean", "LandUse")


############################# 2017: MHW - CLIMATOLOGICALLY AVERAGE SST
#Re-organize 2017 - Temperature
#columns are one for each hour, one for distance2ocean, and one for land use
ReOrg_Temp17_msst <- as.data.frame(matrix(0, ncol=(length(HoursOfInterest) + 2), nrow =n17c))

HourlyMeans17_msst <- HourlyMeans17_msst[order(HourlyMeans17_msst$GridCell),] #make sure the order is consistent so you know which points are which

#Get your hourly temps - response
ReOrg_Temp17_msst[,1] <- HourlyMeans17_msst$X2mTemp[HourlyMeans17_msst$hour == 12]
ReOrg_Temp17_msst[,2] <- HourlyMeans17_msst$X2mTemp[HourlyMeans17_msst$hour == 13]
ReOrg_Temp17_msst[,3] <- HourlyMeans17_msst$X2mTemp[HourlyMeans17_msst$hour == 14]
ReOrg_Temp17_msst[,4] <- HourlyMeans17_msst$X2mTemp[HourlyMeans17_msst$hour == 15]
ReOrg_Temp17_msst[,5] <- HourlyMeans17_msst$X2mTemp[HourlyMeans17_msst$hour == 16]
ReOrg_Temp17_msst[,6] <- HourlyMeans17_msst$X2mTemp[HourlyMeans17_msst$hour == 17]
ReOrg_Temp17_msst[,7] <- HourlyMeans17_msst$X2mTemp[HourlyMeans17_msst$hour == 18]
ReOrg_Temp17_msst[,8] <- HourlyMeans17_msst$X2mTemp[HourlyMeans17_msst$hour == 19]
ReOrg_Temp17_msst[,9] <- HourlyMeans17_msst$X2mTemp[HourlyMeans17_msst$hour == 20]
ReOrg_Temp17_msst[,10] <- HourlyMeans17_msst$X2mTemp[HourlyMeans17_msst$hour == 21]

#Get your fixed effects
ReOrg_Temp17_msst[,11] <- HourlyMeans17_msst$Dist2Ocean[HourlyMeans17_msst$hour == 12]
ReOrg_Temp17_msst[,12] <- HourlyMeans17_msst$LandUse[HourlyMeans17_msst$hour == 12]
names(ReOrg_Temp17_msst) <- c("y.1", "y.2", "y.3", "y.4", "y.5", "y.6", "y.7", "y.8", "y.9", "y.10", "Dist2Ocean", "LandUse")


#re-organize 2017 humidity
ReOrg_RH17_msst <- as.data.frame(matrix(0, ncol=(length(HoursOfInterest) + 2), nrow =n17c))

#Get your hourly temps - response
ReOrg_RH17_msst[,1] <- HourlyMeans17_msst$RelHumid[HourlyMeans17_msst$hour == 12]
ReOrg_RH17_msst[,2] <- HourlyMeans17_msst$RelHumid[HourlyMeans17_msst$hour == 13]
ReOrg_RH17_msst[,3] <- HourlyMeans17_msst$RelHumid[HourlyMeans17_msst$hour == 14]
ReOrg_RH17_msst[,4] <- HourlyMeans17_msst$RelHumid[HourlyMeans17_msst$hour == 15]
ReOrg_RH17_msst[,5] <- HourlyMeans17_msst$RelHumid[HourlyMeans17_msst$hour == 16]
ReOrg_RH17_msst[,6] <- HourlyMeans17_msst$RelHumid[HourlyMeans17_msst$hour == 17]
ReOrg_RH17_msst[,7] <- HourlyMeans17_msst$RelHumid[HourlyMeans17_msst$hour == 18]
ReOrg_RH17_msst[,8] <- HourlyMeans17_msst$RelHumid[HourlyMeans17_msst$hour == 19]
ReOrg_RH17_msst[,9] <- HourlyMeans17_msst$RelHumid[HourlyMeans17_msst$hour == 20]
ReOrg_RH17_msst[,10] <- HourlyMeans17_msst$RelHumid[HourlyMeans17_msst$hour == 21]

#Get your fixed effects
ReOrg_RH17_msst[,11] <- HourlyMeans17_msst$Dist2Ocean[HourlyMeans17_msst$hour == 12]
ReOrg_RH17_msst[,12] <- HourlyMeans17_msst$LandUse[HourlyMeans17_msst$hour == 12]
names(ReOrg_RH17_msst) <- c("y.1", "y.2", "y.3", "y.4", "y.5", "y.6", "y.7", "y.8", "y.9", "y.10", "Dist2Ocean", "LandUse")

```

Next, set-up the priors

Documentation: https://cran.r-project.org/web/packages/spBayes/spBayes.pdf

```{r priors}

#Figure out what parameters you need
N.t <- length(HoursOfInterest) #number of hours we are working with
n.samples <- 2000 #for the MCMC

#these should be the same for 2016 and 2017 but worth checking separately anyway
p16 <- (length(unique(HourlyMeans16$LandUse))) + 1 #number of predictors, have to account for categorical land use
p17 <- (length(unique(HourlyMeans17$LandUse))) + 1 #number of predictors, have to account for categorical land use
p17c <- (length(unique(HourlyMeans17_msst$LandUse))) + 1 #number of predictors, have to account for categorical land use

##starting and priors
x.d <- max(iDist(st_coordinates(coords_UTM)))

#########Need to figure out how to handle relative humditiy because it is bounded between 0 and 100
#########Temperature should be fine

## 2016
tuning16 <- list("phi" = rep(0.1, N.t))

starting16 <-list("beta"=rep(0,N.t*p16), "phi"=rep(3/(0.5*x.d), N.t),
                 "sigma.sq"=rep(1,N.t), "tau.sq"=rep(1, N.t),
                 "sigma.eta"=diag(rep(0.01, p16)))

priors16 <- list("beta.0.Norm"=list(rep(0,p16), diag(1000,p16)),
               "phi.Unif"=list(rep(3/(0.9*x.d),N.t), rep(3/(0.001*x.d),N.t)),
               "sigma.sq.IG"=list(rep(2,N.t), rep(10,N.t)),
               "tau.sq.IG"=list(rep(2,N.t), rep(10,N.t)),
               "sigma.eta.IW"=list(2, diag(0.001,p16)))

## 2017
tuning17 <- list("phi" = rep(0.1, N.t))

starting17 <-list("beta"=rep(0,N.t*p17), "phi"=rep(3/(0.5*x.d), N.t),
                 "sigma.sq"=rep(1,N.t), "tau.sq"=rep(1, N.t),
                 "sigma.eta"=diag(rep(0.01, p17)))

priors17 <- list("beta.0.Norm"=list(rep(0,p17), diag(1000,p17)),
               "phi.Unif"=list(rep(3/(0.9*x.d),N.t), rep(3/(0.001*x.d),N.t)),
               "sigma.sq.IG"=list(rep(2,N.t), rep(10,N.t)),
               "tau.sq.IG"=list(rep(2,N.t), rep(10,N.t)),
               "sigma.eta.IW"=list(2, diag(0.001,p17)))

## 2017
tuning17c <- list("phi" = rep(0.1, N.t))

starting17c <-list("beta"=rep(0,N.t*p17c), "phi"=rep(3/(0.5*x.d), N.t),
                 "sigma.sq"=rep(1,N.t), "tau.sq"=rep(1, N.t),
                 "sigma.eta"=diag(rep(0.01, p17c)))

priors17c <- list("beta.0.Norm"=list(rep(0,p17c), diag(1000,p17c)),
               "phi.Unif"=list(rep(3/(0.9*x.d),N.t), rep(3/(0.001*x.d),N.t)),
               "sigma.sq.IG"=list(rep(2,N.t), rep(10,N.t)),
               "tau.sq.IG"=list(rep(2,N.t), rep(10,N.t)),
               "sigma.eta.IW"=list(2, diag(0.001,p17c)))

```

Now actually run the regression

Temperature should be fine this way, but relative humdity is bounded between 0 and 100. If we don't get close to the ends we may be able to pretend it is approximately continuous but need to check that with the EDA first. 

Update: from EDA seems like you stay between RH 30 and 80, so should be able to approximate it as continuous linear rather than having to deal with a beta regression

You should also decide on the covariance model based on the EDA. Exponential is pretty normal I think...

Update from EDA: Definitely have spatial structure, the difference seems to peak somewhere between 0.5 and 1.0 (distance)

```{r regs 2016}

#Temperature
##get the models
mods <- lapply(paste(c("y.1", "y.2", "y.3", "y.4", "y.5", "y.6", "y.7", "y.8", "y.9", "y.10"),'1 + LandUse + Dist2Ocean',sep='~'), as.formula)

#run the model
m.temp.16 <- spDynLM(mods, data=ReOrg_Temp16, coords=st_coordinates(coords_UTM),
               starting=starting16, tuning=tuning16, priors=priors16, knots = knots,
               cov.model="exponential", n.samples=n.samples, n.report=500) 

#Backup <- m.temp.16

##Relative humidity
m.rh.16 <- spDynLM(mods, data=ReOrg_RH16, coords=st_coordinates(coords_UTM),
               starting=starting16, tuning=tuning16, priors=priors16, knots = knots,
               cov.model="exponential", n.samples=n.samples, n.report=500) 
```


## Look at outputs


```{r 2016 outputs}
#n.samples <- 2000

###############2016 temperature
#Jim's way of calculating
burn.in   <- 0.75*n.samples
sub.samps <- burn.in:n.samples
m.temp.16$p.samples[,"phi"] <- 3/m.temp.16$p.samples[,"phi"]

coeff <- t(apply( m.temp.16$p.beta.samples, 2, quantile, c(.5, .025, .975) ))
se    <- apply( m.temp.16$p.beta.samples, 2, sd )
coeff <- cbind( coeff[,1], se, coeff[,2:3])
print( signif(coeff, 3) )

write.csv(signif(coeff, 3), "../../05AnalysisOutputs/SpDynLM_Coefficients_Temp2016_050225.csv")
saveRDS(m.temp.16, "../../05AnalysisOutputs/SpDynLM_Temp2016_050225.rds")

#spatial random effects
##need to check the order, may be tau then sigma then phi
ts.plot(m.temp.16$p.theta.samples[,1],main="sigma sq",ylab="",
xlim=c(100,nrow(m.temp.16$p.theta.samples)),ylim=c(0,4))
ts.plot(m.temp.16$p.theta.samples[,2],main="tau sq",ylab="",
xlim=c(100,nrow(m.temp.16$p.theta.samples)),ylim=c(0,1))
ts.plot(m.temp.16$p.theta.samples[,3],main="phi",ylab="",
xlim=c(50,nrow(m.temp.16$p.theta.samples)))

##more plots - https://www.rdocumentation.org/packages/spBayes/versions/0.4-7/topics/spDynLM 

coeff <- t(apply( m.temp.16$p.sigma.eta.samples, 2, quantile, c(.5, .025, .975) ))
se    <- apply( m.temp.16$p.sigma.eta.samples, 2, sd )
coeff <- cbind( coeff[,1], se, coeff[,2:3])
print( signif(coeff, 3) )


#########################2016 Relative Humidity
burn.in   <- 0.75*n.samples
sub.samps <- burn.in:n.samples
m.rh.16$p.samples[,"phi"] <- 3/m.rh.16$p.samples[,"phi"]

coeff <- t(apply( m.rh.16$p.beta.samples, 2, quantile, c(.5, .025, .975) ))
se    <- apply( m.rh.16$p.beta.samples, 2, sd )
coeff <- cbind( coeff[,1], se, coeff[,2:3])
print( signif(coeff, 3) )

write.csv(signif(coeff, 3), "../../05AnalysisOutputs/SpDynLM_Coefficients_RelHum2016_050225.csv")
saveRDS(m.rh.16, "../../05AnalysisOutputs/SpDynLM_RelHum2016_050225.rds")
#m.rh.16 <- readRDS("../../05AnalysisOutputs/SpDynLM_RelHum2016_050225.rds")

#spatial random effects
##need to check the order, may be tau then sigma then phi
ts.plot(m.rh.16$p.theta.samples[,1],main="sigma sq",ylab="",
xlim=c(100,nrow(m.rh.16$p.theta.samples)),ylim=c(0,25))
ts.plot(m.rh.16$p.theta.samples[,2],main="tau sq",ylab="",
xlim=c(100,nrow(m.rh.16$p.theta.samples)),ylim=c(0,1.5))
ts.plot(m.rh.16$p.theta.samples[,3],main="phi",ylab="",
xlim=c(50,nrow(m.rh.16$p.theta.samples)))

##more plots - https://www.rdocumentation.org/packages/spBayes/versions/0.4-7/topics/spDynLM 
```

## Run models for MHW year

Now we get into 2017 and do the same as above

```{r regs 2017}

#Temperature
##get the models
mods <- lapply(paste(c("y.1", "y.2", "y.3", "y.4", "y.5", "y.6", "y.7", "y.8", "y.9", "y.10"),'1 + LandUse + Dist2Ocean',sep='~'), as.formula)

#run the model
m.temp.17 <- spDynLM(mods, data=ReOrg_Temp17, coords=st_coordinates(coords_UTM),
               starting=starting17, tuning=tuning17, priors=priors17, knots = knots,
               cov.model="exponential", n.samples=n.samples, n.report=500) 


##Relative humidity
m.rh.17 <- spDynLM(mods, data=ReOrg_RH17, coords=st_coordinates(coords_UTM),
               starting=starting17, tuning=tuning17, priors=priors17, knots = knots,
               cov.model="exponential", n.samples=n.samples, n.report=500) 
```


Now visualize the outputs

```{r 2017 outputs}

###############2017 temperature
#Jim's way of calculating
burn.in   <- 0.75*n.samples
sub.samps <- burn.in:n.samples
m.temp.17$p.samples[,"phi"] <- 3/m.temp.17$p.samples[,"phi"]

coeff <- t(apply( m.temp.17$p.beta.samples, 2, quantile, c(.5, .025, .975) ))
se    <- apply( m.temp.17$p.beta.samples, 2, sd )
coeff <- cbind( coeff[,1], se, coeff[,2:3])
print( signif(coeff, 3) )

write.csv(signif(coeff, 3), "../../05AnalysisOutputs/SpDynLM_Coefficients_Temp2017_050225.csv")
saveRDS(m.temp.17, "../../05AnalysisOutputs/SpDynLM_Temp2017_050225.rds")

#spatial random effects
##need to check the order, may be tau then sigma then phi
ts.plot(m.temp.17$p.theta.samples[,1],main="sigma sq",ylab="",
xlim=c(100,nrow(m.temp.17$p.theta.samples)),ylim=c(0,4))
ts.plot(m.temp.17$p.theta.samples[,2],main="tau sq",ylab="",
xlim=c(100,nrow(m.temp.17$p.theta.samples)),ylim=c(0,1))
ts.plot(m.temp.17$p.theta.samples[,3],main="phi",ylab="",
xlim=c(50,nrow(m.temp.17$p.theta.samples)))

##more plots - https://www.rdocumentation.org/packages/spBayes/versions/0.4-7/topics/spDynLM 

# coeff <- t(apply( m.temp.17$p.sigma.eta.samples, 2, quantile, c(.5, .025, .975) ))
# se    <- apply( m.temp.17$p.sigma.eta.samples, 2, sd )
# coeff <- cbind( coeff[,1], se, coeff[,2:3])
# print( signif(coeff, 3) )


#########################2017 Relative Humidity
burn.in   <- 0.75*n.samples
sub.samps <- burn.in:n.samples
m.rh.17$p.samples[,"phi"] <- 3/m.rh.17$p.samples[,"phi"]

coeff <- t(apply( m.rh.17$p.beta.samples, 2, quantile, c(.5, .025, .975) ))
se    <- apply( m.rh.17$p.beta.samples, 2, sd )
coeff <- cbind( coeff[,1], se, coeff[,2:3])
print( signif(coeff, 3) )

write.csv(signif(coeff, 3), "../../05AnalysisOutputs/SpDynLM_Coefficients_RelHum2017_050225.csv")
saveRDS(m.rh.17, "../../05AnalysisOutputs/SpDynLM_RelHum2017_050225.rds")

#spatial random effects
##need to check the order, may be tau then sigma then phi
ts.plot(m.rh.17$p.theta.samples[,1],main="sigma sq",ylab="",
xlim=c(100,nrow(m.rh.17$p.theta.samples)),ylim=c(0,50))
ts.plot(m.rh.17$p.theta.samples[,2],main="tau sq",ylab="",
xlim=c(100,nrow(m.rh.17$p.theta.samples)),ylim=c(0,1))
ts.plot(m.rh.17$p.theta.samples[,3],main="phi",ylab="",
xlim=c(50,nrow(m.rh.17$p.theta.samples)))

##more plots - https://www.rdocumentation.org/packages/spBayes/versions/0.4-7/topics/spDynLM 

# coeff <- t(apply( m.rh.17$p.sigma.eta.samples, 2, quantile, c(.5, .025, .975) ))
# se    <- apply( m.rh.17$p.sigma.eta.samples, 2, sd )
# coeff <- cbind( coeff[,1], se, coeff[,2:3])
# print( signif(coeff, 3) )

```

#Now do it for simulated 2017
```{r regs 2017 simlated}

#this should work for both datasets
nobs <- length(HourlyMeans17_msst$GridCell) #number of observations
coords <- HourlyMeans17_msst[, c("GridCell", "lat", "long")] #points of the observations
names(coords) <- c("GridCell", "Y", "X") #yes I know X should be longitude, but I somehow flipped it earlier
coords <- unique(coords)

#change your coordinates to UTM so the distances are consistent
coords_sf <- st_as_sf(coords, coords = c("X", "Y"), crs = 4326)
coords_UTM <- st_transform(coords_sf, CRS("+proj=utm +zone=15 ellps=WGS84"))


#Temperature
##get the models
mods <- lapply(paste(c("y.1", "y.2", "y.3", "y.4", "y.5", "y.6", "y.7", "y.8", "y.9", "y.10"),'1 + LandUse + Dist2Ocean',sep='~'), as.formula)

#run the model
m.temp.17c <- spDynLM(mods, data=ReOrg_Temp17_msst, coords=st_coordinates(coords_UTM),
               starting=starting17c, tuning=tuning17c, priors=priors17c, knots = knots,
               cov.model="exponential", n.samples=n.samples, n.report=500) 


##Relative humidity
m.rh.17c <- spDynLM(mods, data=ReOrg_RH17_msst, coords=st_coordinates(coords_UTM),
               starting=starting17c, tuning=tuning17c, priors=priors17c, knots = knots,
               cov.model="exponential", n.samples=n.samples, n.report=500) 
```


Now visualize the outputs

```{r 2017 outputs}

###############2017 temperature
#Jim's way of calculating
burn.in   <- 0.75*n.samples
sub.samps <- burn.in:n.samples
m.temp.17c$p.samples[,"phi"] <- 3/m.temp.17c$p.samples[,"phi"]

coeff <- t(apply( m.temp.17c$p.beta.samples, 2, quantile, c(.5, .025, .975) ))
se    <- apply( m.temp.17c$p.beta.samples, 2, sd )
coeff <- cbind( coeff[,1], se, coeff[,2:3])
print( signif(coeff, 3) )

write.csv(signif(coeff, 3), "../../05AnalysisOutputs/SpDynLM_Coefficients_Temp2017_w16SST_050225.csv")
saveRDS(m.temp.17c, "../../05AnalysisOutputs/SpDynLM_Temp2017_w16SST_050225.rds")

#spatial random effects
##need to check the order, may be tau then sigma then phi
ts.plot(m.temp.17c$p.theta.samples[,1],main="sigma sq",ylab="",
xlim=c(100,nrow(m.temp.17c$p.theta.samples)),ylim=c(0,4))
ts.plot(m.temp.17c$p.theta.samples[,2],main="tau sq",ylab="",
xlim=c(100,nrow(m.temp.17c$p.theta.samples)),ylim=c(0,1))
ts.plot(m.temp.17c$p.theta.samples[,3],main="phi",ylab="",
xlim=c(50,nrow(m.temp.17c$p.theta.samples)))

##more plots - https://www.rdocumentation.org/packages/spBayes/versions/0.4-7/topics/spDynLM 

#########################2017 Relative Humidity
burn.in   <- 0.75*n.samples
sub.samps <- burn.in:n.samples
m.rh.17c$p.samples[,"phi"] <- 3/m.rh.17c$p.samples[,"phi"]

coeff <- t(apply( m.rh.17c$p.beta.samples, 2, quantile, c(.5, .025, .975) ))
se    <- apply( m.rh.17c$p.beta.samples, 2, sd )
coeff <- cbind( coeff[,1], se, coeff[,2:3])
print( signif(coeff, 3) )

write.csv(signif(coeff, 3), "../../05AnalysisOutputs/SpDynLM_Coefficients_RelHum2017_w16SST_050225.csv")
saveRDS(m.rh.17c, "../../05AnalysisOutputs/SpDynLM_RelHum2017_w16SST_050225.rds")
#m.rh.17c <- readRDS('../../05AnalysisOutputs/SpDynLM_RelHum2017_w16SST_050225.rds')

#spatial random effects
##need to check the order, may be tau then sigma then phi
ts.plot(m.rh.17c$p.theta.samples[,1],main="sigma sq",ylab="",
xlim=c(100,nrow(m.rh.17c$p.theta.samples)),ylim=c(0,50))
ts.plot(m.rh.17c$p.theta.samples[,2],main="tau sq",ylab="",
xlim=c(100,nrow(m.rh.17c$p.theta.samples)),ylim=c(0,1))
ts.plot(m.rh.17c$p.theta.samples[,3],main="phi",ylab="",
xlim=c(50,nrow(m.rh.17c$p.theta.samples)))

```

#################Wind speed

```{r get data ready}
#calculate wind speed
HourlyMeans16$WindSpeed <- sqrt(HourlyMeans16$UWind^2 + HourlyMeans16$VWind^2)
HourlyMeans17$WindSpeed <- sqrt(HourlyMeans17$UWind^2 + HourlyMeans17$VWind^2)
HourlyMeans17_msst$WindSpeed <- sqrt(HourlyMeans17_msst$UWind^2 + HourlyMeans17_msst$VWind^2)

#create the storage matrices
ReOrg_WS16 <- as.data.frame(matrix(0, ncol=(length(HoursOfInterest) + 2), nrow = n16))
names(ReOrg_WS16) <- c("y.1", "y.2", "y.3", "y.4", "y.5", "y.6", "y.7", "y.8", "y.9", "y.10", "Dist2Ocean", "LandUse")

ReOrg_WS17 <- as.data.frame(matrix(0, ncol=(length(HoursOfInterest) + 2), nrow = n17))
names(ReOrg_WS17) <- c("y.1", "y.2", "y.3", "y.4", "y.5", "y.6", "y.7", "y.8", "y.9", "y.10", "Dist2Ocean", "LandUse")

ReOrg_WS17s <- as.data.frame(matrix(0, ncol=(length(HoursOfInterest) + 2), nrow = n17c))
names(ReOrg_WS17s) <- c("y.1", "y.2", "y.3", "y.4", "y.5", "y.6", "y.7", "y.8", "y.9", "y.10", "Dist2Ocean", "LandUse")


#Fill the storage matrices - 2016
ReOrg_WS16[,1] <- HourlyMeans16$WindSpeed[HourlyMeans16$hour == 12]
ReOrg_WS16[,2] <- HourlyMeans16$WindSpeed[HourlyMeans16$hour == 13]
ReOrg_WS16[,3] <- HourlyMeans16$WindSpeed[HourlyMeans16$hour == 14]
ReOrg_WS16[,4] <- HourlyMeans16$WindSpeed[HourlyMeans16$hour == 15]
ReOrg_WS16[,5] <- HourlyMeans16$WindSpeed[HourlyMeans16$hour == 16]
ReOrg_WS16[,6] <- HourlyMeans16$WindSpeed[HourlyMeans16$hour == 17]
ReOrg_WS16[,7] <- HourlyMeans16$WindSpeed[HourlyMeans16$hour == 18]
ReOrg_WS16[,8] <- HourlyMeans16$WindSpeed[HourlyMeans16$hour == 19]
ReOrg_WS16[,9] <- HourlyMeans16$WindSpeed[HourlyMeans16$hour == 20]
ReOrg_WS16[,10] <- HourlyMeans16$WindSpeed[HourlyMeans16$hour == 21]
ReOrg_WS16$Dist2Ocean <- HourlyMeans16$Dist2Ocean[HourlyMeans16$hour == 12]
ReOrg_WS16$LandUse <- HourlyMeans16$LandUse[HourlyMeans16$hour == 12]


#Fill the storage matrices - 2017
ReOrg_WS17[,1] <- HourlyMeans17$WindSpeed[HourlyMeans17$hour == 12]
ReOrg_WS17[,2] <- HourlyMeans17$WindSpeed[HourlyMeans17$hour == 13]
ReOrg_WS17[,3] <- HourlyMeans17$WindSpeed[HourlyMeans17$hour == 14]
ReOrg_WS17[,4] <- HourlyMeans17$WindSpeed[HourlyMeans17$hour == 15]
ReOrg_WS17[,5] <- HourlyMeans17$WindSpeed[HourlyMeans17$hour == 16]
ReOrg_WS17[,6] <- HourlyMeans17$WindSpeed[HourlyMeans17$hour == 17]
ReOrg_WS17[,7] <- HourlyMeans17$WindSpeed[HourlyMeans17$hour == 18]
ReOrg_WS17[,8] <- HourlyMeans17$WindSpeed[HourlyMeans17$hour == 19]
ReOrg_WS17[,9] <- HourlyMeans17$WindSpeed[HourlyMeans17$hour == 20]
ReOrg_WS17[,10] <- HourlyMeans17$WindSpeed[HourlyMeans17$hour == 21]
ReOrg_WS17$Dist2Ocean <- HourlyMeans17$Dist2Ocean[HourlyMeans17$hour == 12]
ReOrg_WS17$LandUse <- HourlyMeans17$LandUse[HourlyMeans17$hour == 12]


#Fill the storage matrices - 2017 Simulation
ReOrg_WS17s[,1] <- HourlyMeans17_msst$WindSpeed[HourlyMeans17_msst$hour == 12]
ReOrg_WS17s[,2] <- HourlyMeans17_msst$WindSpeed[HourlyMeans17_msst$hour == 13]
ReOrg_WS17s[,3] <- HourlyMeans17_msst$WindSpeed[HourlyMeans17_msst$hour == 14]
ReOrg_WS17s[,4] <- HourlyMeans17_msst$WindSpeed[HourlyMeans17_msst$hour == 15]
ReOrg_WS17s[,5] <- HourlyMeans17_msst$WindSpeed[HourlyMeans17_msst$hour == 16]
ReOrg_WS17s[,6] <- HourlyMeans17_msst$WindSpeed[HourlyMeans17_msst$hour == 17]
ReOrg_WS17s[,7] <- HourlyMeans17_msst$WindSpeed[HourlyMeans17_msst$hour == 18]
ReOrg_WS17s[,8] <- HourlyMeans17_msst$WindSpeed[HourlyMeans17_msst$hour == 19]
ReOrg_WS17s[,9] <- HourlyMeans17_msst$WindSpeed[HourlyMeans17_msst$hour == 20]
ReOrg_WS17s[,10] <- HourlyMeans17_msst$WindSpeed[HourlyMeans17_msst$hour == 21]
ReOrg_WS17s$Dist2Ocean <- HourlyMeans17_msst$Dist2Ocean[HourlyMeans17_msst$hour == 12]
ReOrg_WS17s$LandUse <- HourlyMeans17_msst$LandUse[HourlyMeans17_msst$hour == 12]
```


```{r regs WS}

#Temperature
##get the models
mods <- lapply(paste(c("y.1", "y.2", "y.3", "y.4", "y.5", "y.6", "y.7", "y.8", "y.9", "y.10"),'1 + LandUse + Dist2Ocean',sep='~'), as.formula)

#run the model
m.ws.16 <- spDynLM(mods, data=ReOrg_WS16, coords=st_coordinates(coords_UTM),
               starting=starting16, tuning=tuning16, priors=priors16, knots = knots,
               cov.model="exponential", n.samples=n.samples, n.report=500) 

m.ws.17 <- spDynLM(mods, data=ReOrg_WS17, coords=st_coordinates(coords_UTM),
               starting=starting17, tuning=tuning17, priors=priors17, knots = knots,
               cov.model="exponential", n.samples=n.samples, n.report=500) 

m.ws.17.s <- spDynLM(mods, data=ReOrg_WS17s, coords=st_coordinates(coords_UTM),
               starting=starting17c, tuning=tuning17c, priors=priors17c, knots = knots,
               cov.model="exponential", n.samples=n.samples, n.report=500) 

```

Now save the outputs

```{r 2017 outputs}

###############2016 wind speed
#Jim's way of calculating
burn.in   <- 0.75*n.samples
sub.samps <- burn.in:n.samples
m.ws.16$p.samples[,"phi"] <- 3/m.ws.16$p.samples[,"phi"]

coeff <- t(apply( m.ws.16$p.beta.samples, 2, quantile, c(.5, .025, .975) ))
se    <- apply( m.ws.16$p.beta.samples, 2, sd )
coeff <- cbind( coeff[,1], se, coeff[,2:3])
print( signif(coeff, 3) )

write.csv(signif(coeff, 3), "../../05AnalysisOutputs/SpDynLM_Coefficients_WindSpeed16_051925.csv")
saveRDS(m.ws.16, "../../05AnalysisOutputs/SpDynLM_WindSpeed16_05192.rds")


###############2017 wind speed
#Jim's way of calculating
burn.in   <- 0.75*n.samples
sub.samps <- burn.in:n.samples
m.ws.17$p.samples[,"phi"] <- 3/m.ws.17$p.samples[,"phi"]

coeff <- t(apply( m.ws.17$p.beta.samples, 2, quantile, c(.5, .025, .975) ))
se    <- apply( m.ws.17$p.beta.samples, 2, sd )
coeff <- cbind( coeff[,1], se, coeff[,2:3])
print( signif(coeff, 3) )

write.csv(signif(coeff, 3), "../../05AnalysisOutputs/SpDynLM_Coefficients_WindSpeed17_051925.csv")
saveRDS(m.ws.17, "../../05AnalysisOutputs/SpDynLM_WindSpeed17_05192.rds")


###############2017 simulated wind speed
#Jim's way of calculating
burn.in   <- 0.75*n.samples
sub.samps <- burn.in:n.samples
m.ws.17.s$p.samples[,"phi"] <- 3/m.ws.17.s$p.samples[,"phi"]

coeff <- t(apply( m.ws.17.s$p.beta.samples, 2, quantile, c(.5, .025, .975) ))
se    <- apply( m.ws.17.s$p.beta.samples, 2, sd )
coeff <- cbind( coeff[,1], se, coeff[,2:3])
print( signif(coeff, 3) )

write.csv(signif(coeff, 3), "../../05AnalysisOutputs/SpDynLM_Coefficients_WindSpeed17_w16SST_051925.csv")
saveRDS(m.ws.17.s, "../../05AnalysisOutputs/SpDynLM_WindSpeed17_w16SST_05192.rds")
```