---
title: "Actors Map Data Restructuring"
author: "Renata Poulton Kamakura"
date: "2024-11-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Look at data

```{r read in data}
DMdata <- read.csv("../01Data/DataModelers.csv")
Stakeholders <- read.csv("../01Data/Stakeholders.csv")

```

## Basic Cleaning - Stakeholders
Have a bunch of rows you don't need
```{r basic clean stakeholders}
##stakeholders
Stakeholders <- Stakeholders[c(!is.null(Stakeholders$Organization.s.) &Stakeholders$Organization.s. !="") ,]

###############Get some useful lists for the tables below

##data modelers

##stakeholders
#get the actual list of stakeholder types
StakeTypeNotes <- Stakeholders$Type.Notes
StakeTypeNotes <- strsplit(StakeTypeNotes, ", ")

#turn it into a list of unique types to reference
sTypesUnique <- c()
for (i in 1:length(StakeTypeNotes)){
  rtypes <- StakeTypeNotes[[i]]
  
  #No type information
  if (length(rtypes) == 0){
    next
  }
  
  #in case you have multiple values within a line
  for (j in 1:length(rtypes)){
    
    #if already in the list
    if (rtypes[j] %in% sTypesUnique){
      next
    
    } else { #if not in the list, add it
      sTypesUnique <- c(sTypesUnique, rtypes[j])
    }
  }
  
}

##Keep track of spelling issues
Incorrect_stk <- c("community", "resilience", "conservation", "Science and Education")
Translate_stk <- c("Community", "Resilience", "Conservation/Restoration", "Education")

##same idea but for locations
StakeLocations <- Stakeholders$Region
StakeLocations <- strsplit(StakeLocations, ", ")

# #turn it into a list of unique types to reference
# sLocUnique <- c()
# for (i in 1:length(StakeLocations)){
#   rlocs <- StakeLocations[[i]]
#   
#   #No location information
#   if (length(rlocs) == 0){
#     next
#   }
#   
#   #in case you have multiple values within a line
#   for (j in 1:length(rlocs)){
#     
#     #if already in the list
#     if (rlocs[j] %in% sLocUnique){
#       next
#     
#     } else { #if not in the list, add it
#       sLocUnique <- c(sLocUnique, rlocs[j])
#     }
#   }
#   
# }

##add states for the ones that are not states
##note that the cities had states attached so don't need to add states for those

#not doing anything with the cities right now, is a layer too complicated of a geography
#CityLocs <- c("Houston", "Galveston", "New Orleans", "St. Petersburg", "Tampa", "Mobile")
GCOOS_sts <- c("TX", "LA", "MS" ,"AL", "FL")
SECOORA_sts <- c("FL", "GA", "SC", "NC")
CARICOOS_sts <- c("PR", "VI")

AllRA_sts <- unique(c(GCOOS_sts, SECOORA_sts, CARICOOS_sts))

##For the future, if something says U.S. or Global I am just going to assign them to all the states.For Gulf of Mexico, going to assign it to GCOOS_sts

```

##Stakeholders
This is the table of unique stakeholder data

```{r stakeholder}
Stk_table <- Stakeholders[,c("Organization.s.", "Contact.Type", "Region", "Contacted", "Last.Contacted")]
Stk_table$SID <- 1:length(Stk_table$Organization.s.)

write.csv(Stk_table, "../01Data/Processed/Stakeholder_Info.csv")
```

## Locations
The main issue here is going to be plotting out the locations. My thought is to split the information into two tables, one that has the unique group data, and one that has a line per state per group. Then you can do a relate with an ID and that can translate between the two but not have a ton of duplicate information. 

Not sure whether Tableau can do this kind of relate in the map, but worth a shot. Once this is set up, it would be easy to create a table that has all the data in one go

```{r locations stakeholder}
##start from StakeLocations
maxRows <- length(AllRA_sts) * length(StakeLocations)

#create the storage table
Loc_table <- as.data.frame(matrix(0, ncol=3, nrow=maxRows))
names(Loc_table) <- c("LID", "SID", "State")

RegionalNames <- c("Global", "U.S.", "Gulf of Mexico")

#go through and fill out the table
rw_iter <- 1
for (i in 1:length(StakeLocations)){
  rlocs <- StakeLocations[[i]]

  #No location information
  if (length(rlocs) == 0){
    print(paste("Warning, no location for", Stakeholders$Organization.s.[i]))
    next
  }

  #in case you have multiple values within a line
  for (j in 1:length(rlocs)){
    
      ##check what the location type is
      #AllRA_sts <- unique(c(GCOOS_sts, SECOORA_sts, CARICOOS_sts))
      if (rlocs[j] %in% AllRA_sts){
        Loc_table$LID[rw_iter] <- rw_iter
        Loc_table$SID[rw_iter] <- Stk_table$SID[i] #get the stakeholder ID
        Loc_table$State[rw_iter] <- rlocs[j]
        
        rw_iter = rw_iter + 1
      } else if (rlocs[j] %in% RegionalNames){ #if it covers the whole area
        
        if (rlocs[j] == "Gulf of Mexico") {#just GCOOS area
          
          #add each Gulf State
          for (l in 1:length(GCOOS_sts)){
            Loc_table$LID[rw_iter] <- rw_iter
            Loc_table$SID[rw_iter] <- Stk_table$SID[i] #get the stakeholder ID
            Loc_table$State[rw_iter] <- GCOOS_sts[l]
            
            rw_iter = rw_iter + 1
          }
          
        } else { #Cover our entire area of interest
          
          ##Add all our states
          for (g in 1:length(AllRA_sts)){
            Loc_table$LID[rw_iter] <- rw_iter
            Loc_table$SID[rw_iter] <- Stk_table$SID[i] #get the stakeholder ID
            Loc_table$State[rw_iter] <- AllRA_sts[g]
            
            rw_iter = rw_iter + 1
          }
          
        } #end if/else for regional groupings
        
      } else{ #skip it because it is presumably a city/state that is not relevant to us rn
        next
      } #end state loop

  } #end overall loop
  
}

#clean it up
Loc_table <- Loc_table[Loc_table$LID !=0,]

write.csv(Loc_table, "../01Data/Processed/Stakeholder_Locs.csv")
```

##Types
This is the table with all the unique type stakeholder combinations
```{r types stakeholder}

##Keep track of spelling issues
Incorrect_stk <- c("community", "resilience", "conservation", "Science and Education", "education")
Translate_stk <- c("Community", "Resilience", "Conservation/Restoration", "Education", "Education")

maxRows <- length(sTypesUnique)*length(Stakeholders$Organization.s.)
Types_tables <- as.data.frame(matrix(0, ncol=3, nrow=maxRows))
names(Types_tables) <- c("TID", "SID", "TypeDescription")

#Go through and add information to the table
rw_iter <- 1
for (i in 1:length(StakeTypeNotes)){
  rtypes <- StakeTypeNotes[[i]]
  
  #No type information
  if (length(rtypes) == 0){
    next
  }
  
  #in case you have multiple values within a line
  for (j in 1:length(rtypes)){
    
    #if need to change the spelling or item
    if (rtypes[j] %in% Incorrect_stk){
      
      rtypes[j] <- Translate_stk[match(rtypes[j], Incorrect_stk)]
    
    }
    
    Types_tables$TID[rw_iter] <- rw_iter
    Types_tables$SID[rw_iter] <- Stk_table$SID[i] #get the stakeholder ID
    Types_tables$TypeDescription[rw_iter] <- rtypes[j]
    
    rw_iter <- rw_iter + 1
    
  } #end of the j loop
  
} #end of the i loop

#clean it up
Types_tables <- Types_tables[Types_tables$TID !=0,]

write.csv(Types_tables, "../01Data/Processed/Stakeholder_Types.csv")

```

##Basic data set up - Data Modelers

Going to need to split out Expertise.Category, Contact.Type, and State.Coverage

```{r data setup data modelers}
##data modelers
DMdata <- DMdata[!is.null(DMdata$Person.Group.Name),]

##for now, going to just say what these are, but could do a similar process as for Stakeholders to get the unique values
u_contactType <- c("Data Collection", "Outreach", "Models")
u_expertiseCat <- c("Ocean observing", "Atmospheric Science", "Stakeholder outreach", "Fisheries", "General observations")

#locations (same as for stakeholders)
GCOOS_sts <- c("TX", "LA", "MS" ,"AL", "FL")
SECOORA_sts <- c("FL", "GA", "SC", "NC")
CARICOOS_sts <- c("PR", "VI")

#for ones that cover the whole region
AllRA_sts <- unique(c(GCOOS_sts, SECOORA_sts, CARICOOS_sts))
```

##Data Modelers

```{r DM basic table}
DM_table <- DMdata[, c("Person.Group.Name","Organization.s.", "Location", "Specific.Location", "Last.Contacted", "Notes")]
DM_table$ID <- 1:length(DM_table$Person.Group.Name)

write.csv(DM_table, "../01Data/Processed/DataModeler_Info.csv")

```