---
title: "MHW Spatial EDA"
author: "Renata Poulton Kamakura"
date: "2025-05-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(sf)
library(tmap)
```

## Input data

Pulling in both the hourly means and the distance to ocean csv files with some simplification of the land use/land cover types to make it more manageable

```{r read in data}

#basic wrfoutput data
HourlyMeans16 <- read.csv("../../03ProcessedData/HourlyMeans_2016_052725.csv")
HourlyMeans17 <- read.csv("../../03ProcessedData/HourlyMeans_2017_052725.csv")
HourlyMeans17_msst <- read.csv("../../03ProcessedData/HourlyMeans_2017_w16SST_052725.csv")

#to simplify the land use types
LU_names = c('Evergreen Needleleaf Forest', 'Evergreen Broadleaf Forest', 
            'Deciduous Needleleaf Forest', 'Deciduous Broadleaf Forest',
            'Mixed Forests', 'Closed Shrublands',  'Open Shrublands', 
            'Woody Savannas', 'Savannas', 'Grasslands', 
            'Permanent wetlands', 'Croplands', 'Urban and Built-Up', 
            'cropland/natural vegetation mosaic', 
            'Snow and Ice', 'Barren or Sparsely Vegetated', 'Water')

NewLUs <- c('Forest', 'Forest', 'Forest', 'Forest', 'Forest',
            'Shrub/Savanna', 'Shrub/Savanna', 'Shrub/Savanna', 'Shrub/Savanna',
            "Shrub/Savanna", "Wetlands", "Crop", "Urban", "Crop", "Snow", 
            "Barren", "Water")
NewLU_nums <- c(1,1,1,1,1,2,2,2,2,2,3,4,5,4,6,7,8)

LU_dict <- as.data.frame(matrix(0, nrow=17, ncol=4))
names(LU_dict) <- c("Original_Name", "Original_Num", "New_Name", "New_Num")

LU_dict$Original_Name <- LU_names
LU_dict$Original_Num <- 1:17
LU_dict$New_Name <- NewLUs
LU_dict$New_Num <- NewLU_nums

#distance from ocean pixels
Dists16 <- read.csv("../../03ProcessedData/Dist2Ocean_2016.csv")
Dists17 <- read.csv("../../03ProcessedData/Dist2Ocean_2017.csv")
Dists17s <- read.csv("../../03ProcessedData/Dist2Ocean_2017_w16SST.csv")

##combine
HourlyMeans16$Dist2Ocean <- Dists16$Dist2Ocean[Dists16$lat == HourlyMeans16$lat & Dists16$long == HourlyMeans16$long]
HourlyMeans17$Dist2Ocean <- Dists17$Dist2Ocean[Dists16$lat == HourlyMeans17$lat & Dists17$long == HourlyMeans17$long]
HourlyMeans17_msst$Dist2Ocean <- Dists17s$Dist2Ocean[Dists17s$lat == HourlyMeans17_msst$lat & Dists17s$long == HourlyMeans17_msst$long]

#########change the hours to local time zone from UTC
HourDict <- as.data.frame(matrix(0, nrow=24, ncol=2))
names(HourDict) <- c("Current", "New")
HourDict$Current <- 0:23
HourDict$New <- c(18, 19, 20, 21, 22, 23, 0, 1, 2, 
                  3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 
                  13, 14, 15, 16, 17)

#convert the hours to CST
HourlyMeans16$hour <- HourDict$New[HourlyMeans16$hour + 1] 
HourlyMeans17$hour <- HourDict$New[HourlyMeans17$hour + 1] 
HourlyMeans17_msst$hour <- HourDict$New[HourlyMeans17_msst$hour + 1] 

#########modify land use
HourlyMeans16_modLU <- HourlyMeans16
HourlyMeans16_modLU$LandUse <- LU_dict$New_Name[HourlyMeans16$LandUse]
HourlyMeans17_modLU <- HourlyMeans17
HourlyMeans17_modLU$LandUse <- LU_dict$New_Name[HourlyMeans17$LandUse]
HourlyMeans17_msst_modLU <- HourlyMeans17_msst
HourlyMeans17_msst_modLU$LandUse <- LU_dict$New_Name[HourlyMeans17_msst$LandUse]

########calculate wind speed
HourlyMeans16_modLU$WindSpeed_10m <- sqrt(HourlyMeans16_modLU$UWind_10m^2 + HourlyMeans16_modLU$VWind_10m^2)
HourlyMeans17_modLU$WindSpeed_10m <- sqrt(HourlyMeans17_modLU$UWind_10m^2 + HourlyMeans17_modLU$VWind_10m^2)
HourlyMeans17_msst_modLU$WindSpeed_10m <- sqrt(HourlyMeans17_msst_modLU$UWind_10m^2 + HourlyMeans17_msst_modLU$VWind_10m^2)

```


#basic visuals of your regions

As background reference to help orient the user to what they will be looking at

```{r basic vis}

# Coordinates
min_lat <- 28.70586
max_lat <- 30.68516
min_lon <- -96.50558
max_lon <- -94.17598
mid_lat <- (min_lat + max_lat) / 2
mid_lon <- (min_lon + max_lon) / 2

# Box creator function
create_box <- function(coords, name) {
  poly <- st_polygon(list(matrix(coords, ncol = 2, byrow = TRUE)))
  st_sf(name = name, geometry = st_sfc(poly, crs = 4326))
}

# Main box
main_coords <- c(
  min_lon, min_lat, #bottom left
  min_lon, max_lat, #top left
  max_lon, max_lat, #top right
  max_lon, min_lat, #bottom right
  min_lon, min_lat #bottom left
)
main_box <- create_box(main_coords, "Overall Region")

# Southwest box
sw_coords <- c(
  -96.2, 28.9, #bottom left
  -96.2, 29.2, #top left
  -95.8, 29.2, #top right
  -95.8, 28.9, #bottom right
  -96.2, 28.9 #bottom left
)
sw_box <- create_box(sw_coords, "Box 1")

# Northeast box
ne_coords <- c(
  -94.8, 30, #bottom left
  -94.8, 30.3, #top left
  -94.4, 30.3, #top right
  -94.4, 30, #bottom right
  -94.8, 30 #bottom left
)
ne_box <- create_box(ne_coords, "Box 3")

# Top-left box
tl_coords <- c(
  -96.4, 30.2, #bottom left
  -96.4, 30.5, #top left
  -96, 30.5, #top right
  -96, 30.2, #bottom right
  -96.4, 30.2 #bottom left
)
tl_box <- create_box(tl_coords, "Box 2")

# Combine all into one sf object
all_boxes <- rbind(main_box, sw_box, ne_box, tl_box)

# Label positions (centroids of each box)
#labels <- st_centroid(all_boxes)

# Plot in view mode
tmap_mode("view")
tm_shape(all_boxes) +
  tm_borders(lwd = 3, col = "blue") +
  tm_basemap("OpenStreetMap") +
  tm_fill("name",
        fill_alpha = 0.4,
        fill.scale = tm_scale(values = "brewer.set2")) + 
  tm_shape(labels) +
  tm_symbols(size = 0) +  # invisible symbol layer
  tm_text("name", size = 1.2, col = "black", options = opt_tm_text(point.label = TRUE))


```

#Now Look at Intersection

Now we want to look at the crop pixels within those three sub-boxes

```{r points within boxes}

#########################create the boxes

# Box 1 - bottom left
b1 <- matrix(c(
  -96.2, 28.9, #bottom left
  -96.2, 29.2, #top left
  -95.8, 29.2, #top right
  -95.8, 28.9, #bottom right
  -96.2, 28.9 #bottom left
), ncol=2, byrow=TRUE)

b1_poly <- st_polygon(list(b1)) |> st_sfc(crs = 4326)

# Box 2 - top left
b2 <- matrix(c(
  -96.4, 30.2, #bottom left
  -96.4, 30.5, #top left
  -96, 30.5, #top right
  -96, 30.2, #bottom right
  -96.4, 30.2 #bottom left
), ncol=2, byrow=TRUE)

b2_poly <- st_polygon(list(b2)) |> st_sfc(crs = 4326)

# Box 3 - top right
b3 <- matrix(c(
  -94.8, 30, #bottom left
  -94.8, 30.3, #top left
  -94.4, 30.3, #top right
  -94.4, 30, #bottom right
  -94.8, 30 #bottom left
), ncol=2, byrow=TRUE)

b3_poly <- st_polygon(list(b3)) |> st_sfc(crs = 4326)


####################create the points

#get the relevant info
pts_16 <- HourlyMeans16_modLU[HourlyMeans16_modLU$hour == 18, c("GridCell", "lat", "long", "LandUse")]
pts_17 <- HourlyMeans17_modLU[HourlyMeans17_modLU$hour == 18, c("GridCell", "lat", "long", "LandUse")]
pts_17s <- HourlyMeans17_msst_modLU[HourlyMeans17_msst_modLU$hour == 18, c("GridCell", "lat", "long", "LandUse")]

#turn them into sf objects
pts_16_sf <- st_as_sf(pts_16, coords = c("long", "lat"), crs = 4326)
pts_17_sf <- st_as_sf(pts_17, coords = c("long", "lat"), crs = 4326)
pts_17s_sf <- st_as_sf(pts_17s, coords = c("long", "lat"), crs = 4326)

#######################look at intersection
#check which are inside - box 1
inside_logical_16b1 <- st_within(pts_16_sf, b1_poly, sparse = FALSE)[, 1]
inside_logical_17b1 <- st_within(pts_17_sf, b1_poly, sparse = FALSE)[, 1]
inside_logical_17sb1 <- st_within(pts_17s_sf, b1_poly, sparse = FALSE)[, 1]

pts_16_inside_b1 <- pts_16_sf[inside_logical_16b1, ]
pts_17_inside_b1 <- pts_17_sf[inside_logical_17b1, ]
pts_17s_inside_b1 <- pts_17s_sf[inside_logical_17sb1, ]

#check which are inside - box 2
inside_logical_16b2 <- st_within(pts_16_sf, b2_poly, sparse = FALSE)[, 1]
inside_logical_17b2 <- st_within(pts_17_sf, b2_poly, sparse = FALSE)[, 1]
inside_logical_17sb2 <- st_within(pts_17s_sf, b2_poly, sparse = FALSE)[, 1]

pts_16_inside_b2 <- pts_16_sf[inside_logical_16b2, ]
pts_17_inside_b2 <- pts_17_sf[inside_logical_17b2, ]
pts_17s_inside_b2 <- pts_17s_sf[inside_logical_17sb2, ]

#check which are inside - box 3
inside_logical_16b3 <- st_within(pts_16_sf, b3_poly, sparse = FALSE)[, 1]
inside_logical_17b3 <- st_within(pts_17_sf, b3_poly, sparse = FALSE)[, 1]
inside_logical_17sb3 <- st_within(pts_17s_sf, b3_poly, sparse = FALSE)[, 1]

pts_16_inside_b3 <- pts_16_sf[inside_logical_16b3, ]
pts_17_inside_b3 <- pts_17_sf[inside_logical_17b3, ]
pts_17s_inside_b3 <- pts_17s_sf[inside_logical_17sb3, ]

```

##Now set up new datasets for analysis

```{r create new df for analysis}

#####################2016
HourlyMeans16_wBoxes <- HourlyMeans16_modLU[HourlyMeans16_modLU$LandUse == "Crop",]
HourlyMeans16_wBoxes$Box <- NA
HourlyMeans16_wBoxes$Box[HourlyMeans16_wBoxes$GridCell %in% pts_16_inside_b1$GridCell] <- 1
HourlyMeans16_wBoxes$Box[HourlyMeans16_wBoxes$GridCell %in% pts_16_inside_b2$GridCell] <- 2
HourlyMeans16_wBoxes$Box[HourlyMeans16_wBoxes$GridCell %in% pts_16_inside_b3$GridCell] <- 3

HourlyMeans16_wBoxes <- HourlyMeans16_wBoxes[!is.na(HourlyMeans16_wBoxes$Box),]

#####################2017
HourlyMeans17_wBoxes <- HourlyMeans17_modLU[HourlyMeans17_modLU$LandUse == "Crop",]
HourlyMeans17_wBoxes$Box <- NA
HourlyMeans17_wBoxes$Box[HourlyMeans17_wBoxes$GridCell %in% pts_17_inside_b1$GridCell] <- 1
HourlyMeans17_wBoxes$Box[HourlyMeans17_wBoxes$GridCell %in% pts_17_inside_b2$GridCell] <- 2
HourlyMeans17_wBoxes$Box[HourlyMeans17_wBoxes$GridCell %in% pts_17_inside_b3$GridCell] <- 3

HourlyMeans17_wBoxes <- HourlyMeans17_wBoxes[!is.na(HourlyMeans17_wBoxes$Box),]


#####################Simulated 2017
HourlyMeans17s_wBoxes <- HourlyMeans17_msst_modLU[HourlyMeans17_msst_modLU$LandUse == "Crop",]
HourlyMeans17s_wBoxes$Box <- NA
HourlyMeans17s_wBoxes$Box[HourlyMeans17s_wBoxes$GridCell %in% pts_17s_inside_b1$GridCell] <- 1
HourlyMeans17s_wBoxes$Box[HourlyMeans17s_wBoxes$GridCell %in% pts_17s_inside_b2$GridCell] <- 2
HourlyMeans17s_wBoxes$Box[HourlyMeans17s_wBoxes$GridCell %in% pts_17s_inside_b3$GridCell] <- 3

HourlyMeans17s_wBoxes <- HourlyMeans17s_wBoxes[!is.na(HourlyMeans17s_wBoxes$Box),]
```

##Basic Plots

Create simple plots of temperature, relative humidity, and wind speed over time for crop pixels within the three boxes

```{r basic plots}
##Box 1 is bottom left
##Box 2 is top left
##Box 3 is top right

###################################Temperature
##2016 Basic Temperature Plot
ggplot(HourlyMeans16_wBoxes, aes(x=as.factor(hour), y=(X2mTemp-273.15), fill=as.factor(Box))) + 
  geom_boxplot() + ggtitle("2016 Temperatures by Region") + 
  xlab("Hour") + ylab("Temperature (C)")  + 
   coord_cartesian(ylim = c(0, 25))

##2017 Basic Temperature Plot
ggplot(HourlyMeans17_wBoxes, aes(x=as.factor(hour), y=(X2mTemp-273.15), fill=as.factor(Box))) + 
  geom_boxplot() + ggtitle("2017 Temperatures by Region") + 
  xlab("Hour") + ylab("Temperature (C)")  + 
   coord_cartesian(ylim = c(0, 25))

##2017 Simulated Basic Temperature Plot
ggplot(HourlyMeans17s_wBoxes, aes(x=as.factor(hour), y=(X2mTemp-273.15), fill=as.factor(Box))) + 
  geom_boxplot() + ggtitle("Simulated 2017 Temperatures by Region") + 
  xlab("Hour") + ylab("Temperature (C)")  + 
   coord_cartesian(ylim = c(0, 25))


###################################Humidity
##2016 Basic Plot
ggplot(HourlyMeans16_wBoxes, aes(x=as.factor(hour), y=(RelHumid), fill=as.factor(Box))) + 
  geom_boxplot() + ggtitle("2016 Relative Humidity by Region") + 
  xlab("Hour") + ylab("Relative Humidity (%)")  + 
   coord_cartesian(ylim = c(0, 100))

##2017 Basic Plot
ggplot(HourlyMeans17_wBoxes, aes(x=as.factor(hour), y=(RelHumid), fill=as.factor(Box))) + 
  geom_boxplot() + ggtitle("2017 Relative Humidity by Region") + 
  xlab("Hour") + ylab("Relative Humidity (%)")  + 
   coord_cartesian(ylim = c(0, 100))

##2017 Simulated Basic Plot
ggplot(HourlyMeans17s_wBoxes, aes(x=as.factor(hour), y=(RelHumid), fill=as.factor(Box))) + 
  geom_boxplot() + ggtitle("Simulated 2017 Relative Humidity by Region") + 
  xlab("Hour") + ylab("Relative Humidity (%)")  + 
   coord_cartesian(ylim = c(0,100))


###################################Wind
##2016 Basic Plot
ggplot(HourlyMeans16_wBoxes, aes(x=as.factor(hour), y=WindSpeed_10m, fill=as.factor(Box))) + 
  geom_boxplot() + ggtitle("2016 Wind Speed by Region") + 
  xlab("Hour") + ylab("Wind Speed (m/s)")  + 
   coord_cartesian(ylim = c(0, 6))

##2017 Basic Plot
ggplot(HourlyMeans17_wBoxes, aes(x=as.factor(hour), y=WindSpeed_10m, fill=as.factor(Box))) + 
  geom_boxplot() + ggtitle("2017 Wind Speed by Region") + 
  xlab("Hour") + ylab("Wind Speed (m/s)")  + 
   coord_cartesian(ylim = c(0, 6))

##2017 Simulated Basic Plot
ggplot(HourlyMeans17s_wBoxes, aes(x=as.factor(hour), y=WindSpeed_10m, fill=as.factor(Box))) + 
  geom_boxplot() + ggtitle("Simulated 2017 Wind Speed by Region") + 
  xlab("Hour") + ylab("Wind Speed (m/s)")  + 
   coord_cartesian(ylim = c(0,6))


```

#Next step
Now I would look at the three boxes comparing MHW to non-MHW both ways

I suspect you will see a bigger difference between the two 2017 runs for box 2, where the wind probably has to go through the city and is disrupted, compared to what you see in the boxes on either side of the city 

Note, all of these are labelled as crop-ish land use

```{r reformat to compare across years}

#for final storage matrix
colnms <- c("Hour", "Box", "Temp16", "Temp17", "Temp17s", "TempDiff17m16", "TempDiff17m17s", 
            "RelHum16", "RelHum17", "RelHum17s", "HumDiff17m16", "HumDiff17m17s",
            "WindSpeed10m16", "WindSpeed10m17", "WindSpeed10m17s", "WSDiff17m16", "WSDiff17m17s")

#Create df
BoxMeans <- as.data.frame(matrix(0, nrow = 24*3, ncol=length(colnms)))
names(BoxMeans) <- colnms

iter <- 1

for (h in c(0:23)){ #hours
  
  for (b in c(1:3)){ #boxes
    
    BoxMeans$Hour[iter] <- h
    BoxMeans$Box[iter] <- b
    
    #############Temperature
    BoxMeans$Temp16[iter] <- mean(HourlyMeans16_wBoxes$X2mTemp[HourlyMeans16_wBoxes$hour == h & HourlyMeans16_wBoxes$Box == b])
    BoxMeans$Temp17[iter] <- mean(HourlyMeans17_wBoxes$X2mTemp[HourlyMeans17_wBoxes$hour == h & HourlyMeans17_wBoxes$Box == b])
    BoxMeans$Temp17s[iter] <- mean(HourlyMeans17s_wBoxes$X2mTemp[HourlyMeans17s_wBoxes$hour == h & HourlyMeans17s_wBoxes$Box == b])
    
    #Calculate the difference
    BoxMeans$TempDiff17m16[iter] <- BoxMeans$Temp17[iter] - BoxMeans$Temp16[iter]
    BoxMeans$TempDiff17m17s[iter] <- BoxMeans$Temp17[iter] - BoxMeans$Temp17s[iter]
    
    ###############Humidity
    BoxMeans$RelHum16[iter] <- mean(HourlyMeans16_wBoxes$RelHumid[HourlyMeans16_wBoxes$hour == h & HourlyMeans16_wBoxes$Box == b])
    BoxMeans$RelHum17[iter] <- mean(HourlyMeans17_wBoxes$RelHumid[HourlyMeans17_wBoxes$hour == h & HourlyMeans17_wBoxes$Box == b])
    BoxMeans$RelHum17s[iter] <- mean(HourlyMeans17s_wBoxes$RelHumid[HourlyMeans17s_wBoxes$hour == h & HourlyMeans17s_wBoxes$Box == b])
    
    #Calculate the difference
    BoxMeans$HumDiff17m16[iter] <- BoxMeans$RelHum17[iter] - BoxMeans$RelHum16[iter]
    BoxMeans$HumDiff17m17s[iter] <- BoxMeans$RelHum17[iter] - BoxMeans$RelHum17s[iter]
    
    
    #################Wind Speed
    BoxMeans$WindSpeed10m16[iter] <- mean(HourlyMeans16_wBoxes$WindSpeed_10m[HourlyMeans16_wBoxes$hour == h & HourlyMeans16_wBoxes$Box == b])
    BoxMeans$WindSpeed10m17[iter] <- mean(HourlyMeans17_wBoxes$WindSpeed_10m[HourlyMeans17_wBoxes$hour == h & HourlyMeans17_wBoxes$Box == b])
    BoxMeans$WindSpeed10m17s[iter] <- mean(HourlyMeans17s_wBoxes$WindSpeed_10m[HourlyMeans17s_wBoxes$hour == h & HourlyMeans17s_wBoxes$Box == b])
    
    #Calculate the difference
    BoxMeans$WSDiff17m16[iter] <- BoxMeans$WindSpeed10m17[iter] - BoxMeans$WindSpeed10m16[iter]
    BoxMeans$WSDiff17m17s[iter] <- BoxMeans$WindSpeed10m17[iter] - BoxMeans$WindSpeed10m17s[iter]
    
    #increase the iterator
    iter <- iter + 1
  }
  
}


```

Look at the difference plots

Plots with header of 2016 are comparing the MHW event to the 2016 non-MHW event. Plots with header of 2017s are comparing the MHW event to the simulated 2017 with 2016 SST non-MHW data. 

```{r diff plots}
##Box 1 is bottom left
##Box 2 is top left
##Box 3 is top right

###################################Temperature
##MHW - 2016
ggplot(BoxMeans, aes(x=as.factor(Hour), y=(TempDiff17m16), group=as.factor(Box))) + 
  geom_line(aes(color=as.factor(Box)), linewidth=1) + ggtitle("MHW - 2016 Temperatures by Region") +  scale_colour_manual('Box Number', values=c("green", "orange", "blue")) +
  xlab("Hour") + ylab("Difference in Temperature (C)")  + 
   coord_cartesian(ylim = c(-1, 6)) +
  theme_minimal()+ geom_hline(yintercept=0, linetype="dashed")


##MHW - 2017s
ggplot(BoxMeans, aes(x=as.factor(Hour), y=(TempDiff17m17s), group=as.factor(Box))) + 
  geom_line(aes(color=as.factor(Box)), linewidth=1) + ggtitle("MHW - Simulated 2017 Temperatures by Region") + scale_colour_manual('Box Number', values=c("green", "orange", "blue")) +
  xlab("Hour") + ylab("Difference in Temperature (C)")  + 
   coord_cartesian(ylim = c(-1, 6)) +
  theme_minimal()+ geom_hline(yintercept=0, linetype="dashed")


###################################Humidity
##MHW - 2016
ggplot(BoxMeans, aes(x=as.factor(Hour), y=(HumDiff17m16), group=as.factor(Box))) + 
  geom_line(aes(color=as.factor(Box)), linewidth=1) + ggtitle("MHW - 2016 Relative Humidity by Region") + scale_colour_manual('Box Number', values=c("green", "orange", "blue")) +
  xlab("Hour") + ylab("Difference in Relative Humidity (%)")  + 
   coord_cartesian(ylim = c(-10, 16)) +
  theme_minimal()+ geom_hline(yintercept=0, linetype="dashed")

##MHW - 2017s
ggplot(BoxMeans, aes(x=as.factor(Hour), y=(HumDiff17m17s), group=as.factor(Box))) + 
  geom_line(aes(color=as.factor(Box)), linewidth=1) + ggtitle("MHW - Simulated 2017 Relative Humidity by Region") + scale_colour_manual('Box Number', values=c("green", "orange", "blue")) +
  xlab("Hour") + ylab("Difference in Relative Humidity (%)")  + 
   coord_cartesian(ylim = c(-10,16)) +
  theme_minimal()+ geom_hline(yintercept=0, linetype="dashed")


###################################Wind
##MHW - 2016
ggplot(BoxMeans, aes(x=as.factor(Hour), y=WSDiff17m16, group=as.factor(Box))) + 
  geom_line(aes(color=as.factor(Box)), linewidth=1) + ggtitle("MHW - 2016 Wind Speed by Region") + scale_colour_manual('Box Number', values=c("green", "orange", "blue")) +
  xlab("Hour") + ylab("Difference in Wind Speed (m/s)")  + 
   coord_cartesian(ylim = c(-0.5, 3.5)) +
  theme_minimal()+ geom_hline(yintercept=0, linetype="dashed")


##MHW - 2017s
ggplot(BoxMeans, aes(x=as.factor(Hour), y=WSDiff17m17s, group=as.factor(Box))) + 
  geom_line(aes(color=as.factor(Box)), linewidth=1) + ggtitle("MHW - Simulated 2017 Wind Speed by Region") + scale_colour_manual('Box Number', values=c("green", "orange", "blue")) +
  xlab("Hour") + ylab("Difference in Wind Speed (m/s)")  + 
   coord_cartesian(ylim = c(-0.5,3.5)) +
  theme_minimal()+ geom_hline(yintercept=0, linetype="dashed")

```