---
title: "WRF Output EDA"
author: "Renata Poulton Kamakura"
date: "2025-04-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(geoR) #variogram functions
```

##Read in data

This is currently an incomplete run

```{r read data}
#Get the hourly averages for each year
#Likely will need to update the files here once re-run with all the data and spin-up
HourlyMeans16 <- read.csv("../../03ProcessedData/HourlyMeans_2016_042225.csv")
HourlyMeans17 <- read.csv("../../03ProcessedData/HourlyMeans_2017_042225.csv")
HourlyMeans17_msst <- read.csv("../../03ProcessedData/HourlyMeans_2017_w16SST_042225.csv")

#More for reference
LU_names = c('Evergreen Needleleaf Forest', 'Evergreen Broadleaf Forest', 
            'Deciduous Needleleaf Forest', 'Deciduous Broadleaf Forest',
            'Mixed Forests', 'Closed Shrublands',  'Open Shrublands', 
            'Woody Savannas', 'Savannas', 'Grasslands', 
            'Permanent wetlands', 'Croplands', 'Urban and Built-Up', 
            'cropland/natural vegetation mosaic', 
            'Snow and Ice', 'Barren or Sparsely Vegetated', 'Water')

NewLUs <- c('Forest', 'Forest', 'Forest', 'Forest', 'Forest',
            'Shrub/Savanna', 'Shrub/Savanna', 'Shrub/Savanna', 'Shrub/Savanna',
            "Shrub/Savanna", "Wetlands", "Crop", "Urban", "Crop", "Snow", 
            "Barren", "Water")
NewLU_nums <- c(1,1,1,1,1,2,2,2,2,2,3,4,5,4,6,7,8)

LU_dict <- as.data.frame(matrix(0, nrow=17, ncol=4))
names(LU_dict) <- c("Original_Name", "Original_Num", "New_Name", "New_Num")

LU_dict$Original_Name <- LU_names
LU_dict$Original_Num <- 1:17
LU_dict$New_Name <- NewLUs
LU_dict$New_Num <- NewLU_nums

```

##Basic EDA

Ideally you want a linear mixed effects model. You can either run it with spatial random effects or use grid cell as a random intercept. My guess is that you should try running it with a spatial random effect but should try to look at the data first 

From looking at this, I suspect the hourly means are not working like they should, these seem too homogenous...

```{r temp EDA}
#A little easier to mess around with
aftHours1 <- c(12,13,14,15,16) #earlier as you get the start of the sea breeze
aftHours2 <- c(17,18,19,20,21) #as the land starts to cool down 

AfternoonMeans16_1 <- HourlyMeans16[HourlyMeans16$hour %in% aftHours1,]
AfternoonMeans17_1 <- HourlyMeans17[HourlyMeans17$hour %in% aftHours1,]
AfternoonMeans17c_1 <- HourlyMeans17_msst[HourlyMeans17_msst$hour %in% aftHours1,]

AfternoonMeans16_2 <- HourlyMeans16[HourlyMeans16$hour %in% aftHours2,]
AfternoonMeans17_2 <- HourlyMeans17[HourlyMeans17$hour %in% aftHours2,]
AfternoonMeans17c_2 <- HourlyMeans17_msst[HourlyMeans17_msst$hour %in% aftHours2,]

##########################temp and lat
ggplot(AfternoonMeans16_1, aes(x=lat, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Early Afternoon Temperatures with Latitude") + 
  xlab("Latitude") + ylab("Temperature (C)")

ggplot(AfternoonMeans16_2, aes(x=lat, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Late Afternoon Temperatures with Latitude") + 
  xlab("Latitude") + ylab("Temperature (C)")



ggplot(AfternoonMeans17_1, aes(x=lat, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Early Afternoon Temperatures with Latitude") + 
  xlab("Latitude") + ylab("Temperature (C)")

ggplot(AfternoonMeans17_2, aes(x=lat, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Late Afternoon Temperatures with Latitude") + 
  xlab("Latitude") + ylab("Temperature (C)")



ggplot(AfternoonMeans17c_1, aes(x=lat, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("Simulated 2017 Early Afternoon Temperatures with Latitude") + 
  xlab("Latitude") + ylab("Temperature (C)")

ggplot(AfternoonMeans17c_2, aes(x=lat, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("Simulated 2017 Late Afternoon Temperatures with Latitude") + 
  xlab("Latitude") + ylab("Temperature (C)")

##########################temp and long
ggplot(AfternoonMeans16_1, aes(x=long, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Early Afternoon Temperatures with Longitude") + 
  xlab("Longitude") + ylab("Temperature (C)")

ggplot(AfternoonMeans16_2, aes(x=long, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Late Afternoon Temperatures with Longitude") + 
  xlab("Longitude") + ylab("Temperature (C)")



ggplot(AfternoonMeans17_1, aes(x=long, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Early Afternoon Temperatures with Longitude") + 
  xlab("Longitude") + ylab("Temperature (C)")

ggplot(AfternoonMeans17_2, aes(x=long, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Late Afternoon Temperatures with Longitude") + 
  xlab("Longitude") + ylab("Temperature (C)")



ggplot(AfternoonMeans17c_1, aes(x=long, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("Simulated 2017 Early Afternoon Temperatures with Longitude") + 
  xlab("Longitude") + ylab("Temperature (C)")

ggplot(AfternoonMeans17c_2, aes(x=long, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("Simulated 2017 Late Afternoon Temperatures with Longitude") + 
  xlab("Longitude") + ylab("Temperature (C)")
```

```{r figs section 2}
############################temp and land use + hour
AfternoonMeans16_modLU_1 <- AfternoonMeans16_1
AfternoonMeans16_modLU_1$LandUse <- LU_dict$New_Name[AfternoonMeans16_1$LandUse]
AfternoonMeans16_modLU_2 <- AfternoonMeans16_2
AfternoonMeans16_modLU_2$LandUse<- LU_dict$New_Name[AfternoonMeans16_2$LandUse]

AfternoonMeans17_modLU_1 <- AfternoonMeans17_1
AfternoonMeans17_modLU_1$LandUse <- LU_dict$New_Name[AfternoonMeans17_1$LandUse]
AfternoonMeans17_modLU_2 <- AfternoonMeans17_2
AfternoonMeans17_modLU_2$LandUse <- LU_dict$New_Name[AfternoonMeans17_2$LandUse] 

AfternoonMeans17c_modLU_1 <- AfternoonMeans17c_1
AfternoonMeans17c_modLU_1$LandUse <- LU_dict$New_Name[AfternoonMeans17c_1$LandUse]
AfternoonMeans17c_modLU_2 <- AfternoonMeans17c_2
AfternoonMeans17c_modLU_2$LandUse <- LU_dict$New_Name[AfternoonMeans17c_2$LandUse] 

##ideally should group land use somehow
ggplot(AfternoonMeans16_modLU_1, aes(x=as.factor(hour), y=(X2mTemp-273.15), fill=LandUse)) + 
  geom_boxplot() + ggtitle("2016 Early Afternoon Temperatures by Land Use") + 
  xlab("Hour") + ylab("Temperature (C)")  + 
   coord_cartesian(ylim = c(0, 25))

ggplot(AfternoonMeans16_modLU_2, aes(x=as.factor(hour), y=(X2mTemp-273.15), fill=LandUse)) + 
  geom_boxplot() + ggtitle("2016 Late Afternoon Temperatures by Land Use") + 
  xlab("Hour") + ylab("Temperature (C)")+ 
   coord_cartesian(ylim = c(0, 25))


ggplot(AfternoonMeans17_modLU_1, aes(x=as.factor(hour), y=(X2mTemp-273.15), fill=LandUse)) + 
  geom_boxplot() + ggtitle("2017 Early Afternoon Temperatures by Land Use") + 
  xlab("Hour") + ylab("Temperature (C)")+ 
   coord_cartesian(ylim = c(0, 25))

ggplot(AfternoonMeans17_modLU_2, aes(x=as.factor(hour), y=(X2mTemp-273.15), fill=LandUse)) + 
  geom_boxplot() + ggtitle("2017 Late Afternoon Temperatures by Land Use") + 
  xlab("Hour") + ylab("Temperature (C)")+ 
   coord_cartesian(ylim = c(0, 25))


ggplot(AfternoonMeans17c_modLU_1, aes(x=as.factor(hour), y=(X2mTemp-273.15), fill=LandUse)) + 
  geom_boxplot() + ggtitle("Simulated 2017 Early Afternoon Temperatures by Land Use") + 
  xlab("Hour") + ylab("Temperature (C)")+ 
   coord_cartesian(ylim = c(0, 25))

ggplot(AfternoonMeans17c_modLU_2, aes(x=as.factor(hour), y=(X2mTemp-273.15), fill=LandUse)) + 
  geom_boxplot() + ggtitle("Simulated 2017 Late Afternoon Temperatures by Land Use") + 
  xlab("Hour") + ylab("Temperature (C)")+ 
   coord_cartesian(ylim = c(0, 25))

############################temp and distance to ocean

ggplot(AfternoonMeans16_1, aes(x=Dist2Ocean, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Early Afternoon Temperatures with Distance to Ocean") + 
  xlab("Distance to Ocean (m)") + ylab("Temperature (C)")

ggplot(AfternoonMeans16_2, aes(x=Dist2Ocean, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Late Afternoon Temperatures with Distance to Ocean") + 
  xlab("Distance to Ocean (m)") + ylab("Temperature (C)")



ggplot(AfternoonMeans17_1, aes(x=Dist2Ocean, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Early Afternoon Temperatures with Distance to Ocean") + 
  xlab("Distance to Ocean (m)") + ylab("Temperature (C)")

ggplot(AfternoonMeans17_2, aes(x=Dist2Ocean, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Late Afternoon Temperatures with Distance to Ocean") + 
  xlab("Distance to Ocean (m)") + ylab("Temperature (C)")



ggplot(AfternoonMeans17c_1, aes(x=Dist2Ocean, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("Simulated 2017 Early Afternoon Temperatures with Distance to Ocean") + 
  xlab("Distance to Ocean (m)") + ylab("Temperature (C)")

ggplot(AfternoonMeans17c_2, aes(x=Dist2Ocean, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("Simulated 2017 Late Afternoon Temperatures with Distance to Ocean") + 
  xlab("Distance to Ocean (m)") + ylab("Temperature (C)")
```


```{r humidity eda}

####################Relative Humidity


###########################rel humidity and lat
ggplot(AfternoonMeans16_1, aes(x=lat, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Early Afternoon Relative Humidity with Latitude") + 
  xlab("Latitude") + ylab("Relative Humidity (%)")

ggplot(AfternoonMeans16_2, aes(x=lat, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Late Afternoon Relative Humidity with Latitude") + 
  xlab("Latitude") + ylab("Relative Humidity (%)")


ggplot(AfternoonMeans17_1, aes(x=lat, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Early Afternoon Relative Humidity with Latitude") + 
  xlab("Latitude") + ylab("Relative Humidity (%)")

ggplot(AfternoonMeans17_2, aes(x=lat, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Late Afternoon Relative Humidity with Latitude") + 
  xlab("Latitude") + ylab("Relative Humidity (%)")


ggplot(AfternoonMeans17c_1, aes(x=lat, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("Simulated 2017 Early Afternoon Relative Humidity with Latitude") + 
  xlab("Latitude") + ylab("Relative Humidity (%)")

ggplot(AfternoonMeans17c_2, aes(x=lat, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("Simulated 2017 Late Afternoon Relative Humidity with Latitude") + 
  xlab("Latitude") + ylab("Relative Humidity (%)")



#########################rel humidity and long
ggplot(AfternoonMeans16_1, aes(x=long, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Early Afternoon Relative Humidity with Longitude") + 
  xlab("Longitude") + ylab("Relative Humidity (%)")

ggplot(AfternoonMeans16_2, aes(x=long, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Late Afternoon Relative Humidity with Longitude") + 
  xlab("Longitude") + ylab("Relative Humidity (%)")

ggplot(AfternoonMeans17_1, aes(x=long, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Early Afternoon Relative Humidity with Longitude") + 
  xlab("Longitude") + ylab("Relative Humidity (%)")

ggplot(AfternoonMeans17_2, aes(x=long, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Late Afternoon Relative Humidity with Longitude") + 
  xlab("Longitude") + ylab("Relative Humidity (%)")


ggplot(AfternoonMeans17c_1, aes(x=long, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("Simulated 2017 Early Afternoon Relative Humidity with Longitude") + 
  xlab("Longitude") + ylab("Relative Humidity (%)")

ggplot(AfternoonMeans17c_2, aes(x=long, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("Simulated 2017 Late Afternoon Relative Humidity with Longitude") + 
  xlab("Longitude") + ylab("Relative Humidity (%)")

##################rel humidity and land use
ggplot(AfternoonMeans16_modLU_1, aes(x=as.factor(hour), y=RelHumid, fill=LandUse)) + 
  geom_boxplot() + ggtitle("2016 Early Afternoon Relative Humidity by Land Use") + 
  xlab("Hour") + ylab("Relative Humidity (%)") + 
   coord_cartesian(ylim = c(0, 100))

ggplot(AfternoonMeans16_modLU_2, aes(x=as.factor(hour), y=RelHumid, fill=LandUse)) + 
  geom_boxplot() + ggtitle("2016 Late Afternoon Relative Humidity by Land Use") + 
  xlab("Hour") + ylab("Relative Humidity (%)") + 
   coord_cartesian(ylim = c(0, 100))


ggplot(AfternoonMeans17_modLU_1, aes(x=as.factor(hour), y=RelHumid, fill=LandUse)) + 
  geom_boxplot() + ggtitle("2017 Early Afternoon Relative Humidity by Land Use") + 
  xlab("Hour") + ylab("Relative Humidity (%)") + 
   coord_cartesian(ylim = c(0, 100))

ggplot(AfternoonMeans17_modLU_2, aes(x=as.factor(hour), y=RelHumid, fill=LandUse)) + 
  geom_boxplot() + ggtitle("2017 Late Afternoon Relative Humidity by Land Use") + 
  xlab("Hour") + ylab("Relative Humidity (%)") + 
   coord_cartesian(ylim = c(0, 100))



ggplot(AfternoonMeans17c_modLU_1, aes(x=as.factor(hour), y=RelHumid, fill=LandUse)) + 
  geom_boxplot() + ggtitle("Simulated 2017 Early Afternoon Relative Humidity by Land Use") + 
  xlab("Hour") + ylab("Relative Humidity (%)") + 
   coord_cartesian(ylim = c(0, 100))

ggplot(AfternoonMeans17c_modLU_2, aes(x=as.factor(hour), y=RelHumid, fill=LandUse)) + 
  geom_boxplot() + ggtitle("Simulated 2017 Late Afternoon Relative Humidity by Land Use") + 
  xlab("Hour") + ylab("Relative Humidity (%)") + 
   coord_cartesian(ylim = c(0, 100))


#########################rel humidity and distance to ocean
ggplot(AfternoonMeans16_1, aes(x=Dist2Ocean, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Early Afternoon Relative Humidity with Distance to Ocean") + 
  xlab("Distance to Ocean (m)") + ylab("Relative Humidity (%)")

ggplot(AfternoonMeans16_2, aes(x=Dist2Ocean, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Late Afternoon Relative Humidity with Distance to Ocean") + 
  xlab("Distance to Ocean (m)") + ylab("Relative Humidity (%)")



ggplot(AfternoonMeans17_1, aes(x=Dist2Ocean, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Early Afternoon Relative Humidity with Distance to Ocean") + 
  xlab("Distance to Ocean (m)") + ylab("Relative Humidity (%)")

ggplot(AfternoonMeans17_2, aes(x=Dist2Ocean, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Late Afternoon Relative Humidity with Distance to Ocean") + 
  xlab("Distance to Ocean (m)") + ylab("Relative Humidity (%)")



ggplot(AfternoonMeans17c_1, aes(x=Dist2Ocean, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("Simulated 2017 Early Afternoon Relative Humidity with Distance to Ocean") + 
  xlab("Distance to Ocean (m)") + ylab("Relative Humidity (%)")

ggplot(AfternoonMeans17c_2, aes(x=Dist2Ocean, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("Simulated 2017 Late Afternoon Relative Humidity with Distance to Ocean") + 
  xlab("Distance to Ocean (m)") + ylab("Relative Humidity (%)")

```

##EDA - variogram

See if the spatial random effects approach is worth it

```{r variogram temp}
#Pick an hour of data to look at to make this a bit simpler
#thinking 17 since it is kind of between the shift in aft to evening temps
MidAftTemp16 <- HourlyMeans16[HourlyMeans16$hour == 17,]
MidAftTemp17 <- HourlyMeans17[HourlyMeans17$hour == 17,]
MidAftTemp17c <- HourlyMeans17_msst[HourlyMeans17_msst$hour == 17,]

##start to look at varigra

ms for temp - 2016
varDatTemp16 <- as.geodata(MidAftTemp16, coords.col = 2:3, data.col=5)
plot(varDatTemp16)

# v1Temp16 <- variog(varDatTemp16, trend="1st", option="bin")
# plot(v1Temp16)
# #v1Temp16.c <- variog(varDatTemp16, op="cloud")
# v1Temp16.bc <- variog(varDatTemp16, bin.cloud=TRUE)

#plot(v1Temp16.c, main="variogram cloud 2016 Temp")
plot(v1Temp16.bc, bin.cloud=TRUE, main="clouds for binned variogram 2016 Temp")

##variogram for temp 2017
varDatTemp17 <- as.geodata(MidAftTemp17, coords.col = 2:3, data.col=5)
plot(varDatTemp17)

# v1Temp17 <- variog(varDatTemp17, trend="1st", option="bin")
# plot(v1Temp17)
# #v1Temp17.c <- variog(varDatTemp17, op="cloud")
# v1Temp17.bc <- variog(varDatTemp17, bin.cloud=TRUE)

#plot(v1Temp17.c, main="variogram cloud 2017 Temp")
plot(v1Temp17.bc, bin.cloud=TRUE, main="clouds for binned variogram 2017 Temp")

##variogram for simulated temp 2017
varDatTemp17c <- as.geodata(MidAftTemp17c, coords.col = 2:3, data.col=5)
plot(varDatTemp17)

# v1Temp17c <- variog(varDatTemp17c, trend="1st", option="bin")
# plot(v1Temp17c)
# v1Temp17c.bc <- variog(varDatTemp17c, bin.cloud=TRUE)
# 
# plot(v1Temp17c.bc, bin.cloud=TRUE, main="clouds for binned variogram simulated 2017 Temp")

```


```{r variogram rh}

#variograms for RH - 2016
varDatRH16 <- as.geodata(MidAftTemp16, coords.col = 2:3, data.col=6)
plot(varDatRH16)

# v1RH16 <- variog(varDatRH16, trend="1st", option="bin")
# plot(v1RH16)
# #v1RH16.c <- variog(varDatRH16, op="cloud")
# v1RH16.bc <- variog(varDatRH16, bin.cloud=TRUE)

#plot(v1RH16.c, main="variogram cloud 2016 Relative Humidity")
#plot(v1RH16.bc, bin.cloud=TRUE, main="clouds for binned variogram 2016 Relative Humidity")

#varigrams for RH 2017
varDatRH17 <- as.geodata(MidAftTemp17, coords.col = 2:3, data.col=6)
plot(varDatRH17)

# v1RH17 <- variog(varDatRH17, trend="1st", option="bin")
# plot(v1RH17)
# #v1RH17.c <- variog(varDatRH17, op="cloud")
# v1RH17.bc <- variog(varDatRH17, bin.cloud=TRUE)

#plot(v1RH17.c, main="variogram cloud 2017 Relative Humidity")
#plot(v1RH17.bc, bin.cloud=TRUE, main="clouds for binned variogram 2017 Relative Humidity")

#varigrams for RH 2017
varDatRH17c <- as.geodata(MidAftTemp17c, coords.col = 2:3, data.col=6)
plot(varDatRH17c)

# v1RH17c <- variog(varDatRH17c, trend="1st", option="bin")
# plot(v1RH17c)
# v1RH17c.bc <- variog(varDatRH17c, bin.cloud=TRUE)

#plot(v1RH17c.bc, bin.cloud=TRUE, main="clouds for binned variogram 2017 simulated Relative Humidity")

```

#EDA Temp differences

Just go through and do basic differences by LU type

Note: since I am averaging, this is very coarse, it reduces all the variation within land use types

```{r means by LU}
HourlyMeans16_modLU <- HourlyMeans16
HourlyMeans16_modLU$LandUse <- LU_dict$New_Name[HourlyMeans16$LandUse]

HourlyMeans17_modLU <- HourlyMeans17
HourlyMeans17_modLU$LandUse <- LU_dict$New_Name[HourlyMeans17$LandUse]

HourlyMeans17_msst_modLU <- HourlyMeans17_msst
HourlyMeans17_msst_modLU$LandUse <- LU_dict$New_Name[HourlyMeans17_msst$LandUse]

#create storage matrix
colNms <- c("Hour", "LandUse", "MeanRH16", "MeanRH17", "MeanRH17s", 
            "MeanTemp16", "MeanTemp17", "MeanTemp17s")
uLU <- unique(LU_dict$New_Name)
uLUnm <- unique(LU_dict$New_Num)
HourlyMean_byLU <- as.data.frame(matrix(0, ncol=length(colNms), nrow=length(uLU)*24))
names(HourlyMean_byLU) <- colNms

#set up your index for the storage matrix
smat_ind <- 1

for (h in 0:23){
  
  #get hourly data
  hr16 <- HourlyMeans16_modLU[HourlyMeans16_modLU$hour == h, ]
  hr17 <- HourlyMeans17_modLU[HourlyMeans17_modLU$hour == h, ]
  hr17s <- HourlyMeans17_msst_modLU[HourlyMeans17_msst_modLU$hour == h, ]
  
  for (l in 1:length(uLUnm)){
    
    HourlyMean_byLU$Hour[smat_ind] <- h
    HourlyMean_byLU$LandUse[smat_ind] <- uLU[l]
    HourlyMean_byLU$MeanRH16[smat_ind] <- mean(hr16$RelHumid[hr16$LandUse == uLU[l]])
    HourlyMean_byLU$MeanRH17[smat_ind] <- mean(hr17$RelHumid[hr17$LandUse == uLU[l]])
    HourlyMean_byLU$MeanRH17s[smat_ind] <- mean(hr17s$RelHumid[hr17s$LandUse == uLU[l]])
    HourlyMean_byLU$MeanTemp16[smat_ind] <- mean(hr16$X2mTemp[hr16$LandUse == uLU[l]])
    HourlyMean_byLU$MeanTemp17[smat_ind] <- mean(hr17$X2mTemp[hr17$LandUse == uLU[l]])
    HourlyMean_byLU$MeanTemp17s[smat_ind] <- mean(hr17s$X2mTemp[hr17s$LandUse == uLU[l]])
    
    #increase the iteration
    smat_ind <- smat_ind + 1
  }
  
}

#get rid of LU with no data
HourlyMean_byLU <- HourlyMean_byLU[!is.na(HourlyMean_byLU$MeanTemp16),]

#calculate your differences
HourlyMean_byLU$DiffRH17m16 <- HourlyMean_byLU$MeanRH17 - HourlyMean_byLU$MeanRH16
HourlyMean_byLU$DiffRH17m17s <- HourlyMean_byLU$MeanRH17 - HourlyMean_byLU$MeanRH17s
HourlyMean_byLU$DiffTemp17m16 <- HourlyMean_byLU$MeanTemp17 - HourlyMean_byLU$MeanTemp16
HourlyMean_byLU$DiffTemp17m17s <- HourlyMean_byLU$MeanTemp17 - HourlyMean_byLU$MeanTemp17s
```


Now plot out our results so we can look at it a bit more clearly

```{r diff visuals temp}

###TEMPERATURE
ggplot(HourlyMean_byLU, aes(x=as.factor(Hour), y=(DiffTemp17m16), group=LandUse)) + 
  geom_line(aes(color=LandUse), size=1) + ggtitle("2017 Minus 2016 Temperatures by Land Use") + 
  xlab("Hour") + ylab("Temperature Difference (C)")  + geom_hline(yintercept=0, linetype="dashed") + 
   coord_cartesian(ylim = c(-1, 6))

ggplot(HourlyMean_byLU, aes(x=as.factor(Hour), y=(DiffTemp17m17s), group=LandUse)) + 
  geom_line(aes(color=LandUse), size=1) + ggtitle("2017 Minus Simulated 2017 Temperatures by Land Use") + 
  xlab("Hour") + ylab("Temperature Difference (C)")  + geom_hline(yintercept=0, linetype="dashed") + 
   coord_cartesian(ylim = c(-1, 6))


####to figure out what is happening at 1pm

ggplot(HourlyMean_byLU, aes(x=as.factor(Hour), y=(MeanTemp16-273.15), group=LandUse)) + 
  geom_line(aes(color=LandUse), size=1) + ggtitle("2016 Temperatures by Land Use") + 
  xlab("Hour") + ylab("Temperature (C)") + 
   coord_cartesian(ylim = c(4, 24))

ggplot(HourlyMean_byLU, aes(x=as.factor(Hour), y=(MeanTemp17-273.15), group=LandUse)) + 
  geom_line(aes(color=LandUse), size=1) + ggtitle("2017 Temperatures by Land Use") + 
  xlab("Hour") + ylab("Temperature (C)") + 
   coord_cartesian(ylim = c(4, 24))

ggplot(HourlyMean_byLU, aes(x=as.factor(Hour), y=(MeanTemp17s-273.15), group=LandUse)) + 
  geom_line(aes(color=LandUse), size=1) + ggtitle("Simulated 2017 Temperatures by Land Use") + 
  xlab("Hour") + ylab("Temperature (C)") + 
   coord_cartesian(ylim = c(4, 24))

```

Now for relative humidity

```{r diff visuals RH}

ggplot(HourlyMean_byLU, aes(x=as.factor(Hour), y=(DiffRH17m16), group=LandUse)) + 
  geom_line(aes(color=LandUse), size=1) + ggtitle("2017 Minus 2016 Relative Humidity by Land Use") + 
  xlab("Hour") + ylab("RH difference Difference (%)")  + geom_hline(yintercept=0, linetype="dashed") + 
   coord_cartesian(ylim = c(-10, 18))

ggplot(HourlyMean_byLU, aes(x=as.factor(Hour), y=(DiffRH17m17s), group=LandUse)) + 
  geom_line(aes(color=LandUse), size=1) + ggtitle("2017 Minus Simulated 2017 Relative Humidity by Land Use") + 
  xlab("Hour") + ylab("RH difference Difference (%)") + geom_hline(yintercept=0, linetype="dashed")  + 
   coord_cartesian(ylim = c(-10, 18))

###to see them each more clearly
ggplot(HourlyMean_byLU, aes(x=as.factor(Hour), y=(MeanRH16), group=LandUse)) + 
  geom_line(aes(color=LandUse), size=1) + ggtitle("2016 Relative Humidity by Land Use") + 
  xlab("Hour") + ylab("RH (%)") + geom_hline(yintercept=0, linetype="dashed")  + 
   coord_cartesian(ylim = c(0, 100))

ggplot(HourlyMean_byLU, aes(x=as.factor(Hour), y=(MeanRH17), group=LandUse)) + 
  geom_line(aes(color=LandUse), size=1) + ggtitle("2017 Relative Humidity by Land Use") + 
  xlab("Hour") + ylab("RH (%)") + geom_hline(yintercept=0, linetype="dashed")  + 
   coord_cartesian(ylim = c(0, 100))

ggplot(HourlyMean_byLU, aes(x=as.factor(Hour), y=(MeanRH17s), group=LandUse)) + 
  geom_line(aes(color=LandUse), size=1) + ggtitle("Simulated 2017 Relative Humidity by Land Use") + 
  xlab("Hour") + ylab("RH (%)") + geom_hline(yintercept=0, linetype="dashed")  + 
   coord_cartesian(ylim = c(0, 100))

```