---
title: "MHW Spatial EDA"
author: "Renata Poulton Kamakura"
date: "2025-05-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(sf)
library(tmap)
library(dplyr)
```

## Input data

Pulling in both the hourly means and the distance to ocean csv files with some simplification of the land use/land cover types to make it more manageable

```{r read in data}

#basic wrfoutput data
HourlyMeans16_0 <- read.csv("../../03ProcessedData/HourlyMeans_2016_0.csv")
HourlyMeans16_1 <- read.csv("../../03ProcessedData/HourlyMeans_2016_1.csv")
HourlyMeans16_2 <- read.csv("../../03ProcessedData/HourlyMeans_2016_2.csv")

HourlyMeans17_0 <- read.csv("../../03ProcessedData/HourlyMeans_2017_0.csv")
HourlyMeans17_1 <- read.csv("../../03ProcessedData/HourlyMeans_2017_1.csv")
HourlyMeans17_2 <- read.csv("../../03ProcessedData/HourlyMeans_2017_1.csv")

#to simplify the land use types
LU_names = c('Evergreen Needleleaf Forest', 'Evergreen Broadleaf Forest', 
            'Deciduous Needleleaf Forest', 'Deciduous Broadleaf Forest',
            'Mixed Forests', 'Closed Shrublands',  'Open Shrublands', 
            'Woody Savannas', 'Savannas', 'Grasslands', 
            'Permanent wetlands', 'Croplands', 'Urban and Built-Up', 
            'cropland/natural vegetation mosaic', 
            'Snow and Ice', 'Barren or Sparsely Vegetated', 'Water')

NewLUs <- c('Forest', 'Forest', 'Forest', 'Forest', 'Forest',
            'Shrub/Savanna', 'Shrub/Savanna', 'Shrub/Savanna', 'Shrub/Savanna',
            "Shrub/Savanna", "Wetlands", "Crop", "Urban", "Crop", "Snow", 
            "Barren", "Water")
NewLU_nums <- c(1,1,1,1,1,2,2,2,2,2,3,4,5,4,6,7,8)

LU_dict <- as.data.frame(matrix(0, nrow=17, ncol=4))
names(LU_dict) <- c("Original_Name", "Original_Num", "New_Name", "New_Num")

LU_dict$Original_Name <- LU_names
LU_dict$Original_Num <- 1:17
LU_dict$New_Name <- NewLUs
LU_dict$New_Num <- NewLU_nums

#distance from ocean pixels
Dists16 <- read.csv("../../03ProcessedData/Dist2Ocean_2016_0.csv")
Dists17 <- read.csv("../../03ProcessedData/Dist2Ocean_2017_0.csv")

##combine
HourlyMeans16_0$Dist2Ocean <- Dists16$Dist2Ocean[Dists16$lat == HourlyMeans16_0$lat & Dists16$long == HourlyMeans16_0$lon]
HourlyMeans16_1$Dist2Ocean <- Dists16$Dist2Ocean[Dists16$lat == HourlyMeans16_1$lat & Dists16$long == HourlyMeans16_1$lon]
HourlyMeans16_2$Dist2Ocean <- Dists16$Dist2Ocean[Dists16$lat == HourlyMeans16_2$lat & Dists16$long == HourlyMeans16_2$lon]

HourlyMeans17_0$Dist2Ocean <- Dists17$Dist2Ocean[Dists16$lat == HourlyMeans17_0$lat & Dists17$long == HourlyMeans17_0$lon]
HourlyMeans17_1$Dist2Ocean <- Dists17$Dist2Ocean[Dists16$lat == HourlyMeans17_1$lat & Dists17$long == HourlyMeans17_1$lon]
HourlyMeans17_2$Dist2Ocean <- Dists17$Dist2Ocean[Dists16$lat == HourlyMeans17_2$lat & Dists17$long == HourlyMeans17_2$lon]


#########change the hours to local time zone from UTC
HourDict <- as.data.frame(matrix(0, nrow=24, ncol=2))
names(HourDict) <- c("Current", "New")
HourDict$Current <- 0:23
HourDict$New <- c(18, 19, 20, 21, 22, 23, 0, 1, 2, 
                  3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 
                  13, 14, 15, 16, 17)

#convert the hours to CST
HourlyMeans16_0$hour <- HourDict$New[HourlyMeans16_0$hour + 1] 
HourlyMeans16_1$hour <- HourDict$New[HourlyMeans16_1$hour + 1] 
HourlyMeans16_2$hour <- HourDict$New[HourlyMeans16_2$hour + 1] 

HourlyMeans17_0$hour <- HourDict$New[HourlyMeans17_0$hour + 1] 
HourlyMeans17_1$hour <- HourDict$New[HourlyMeans17_1$hour + 1] 
HourlyMeans17_2$hour <- HourDict$New[HourlyMeans17_2$hour + 1] 

#########modify land use
HourlyMeans16_0$landuse <- LU_dict$New_Name[HourlyMeans16_0$landuse]
HourlyMeans16_1$landuse <- LU_dict$New_Name[HourlyMeans16_1$landuse]
HourlyMeans16_2$landuse <- LU_dict$New_Name[HourlyMeans16_2$landuse]

HourlyMeans17_0$landuse <- LU_dict$New_Name[HourlyMeans17_0$landuse]
HourlyMeans17_1$landuse <- LU_dict$New_Name[HourlyMeans17_1$landuse]
HourlyMeans17_2$landuse <- LU_dict$New_Name[HourlyMeans17_2$landuse]

```


#basic visuals of your regions

As background reference to help orient the user to what they will be looking at

```{r basic vis}

# Coordinates
min_lat <- 28.70586
max_lat <- 30.68516
min_lon <- -96.50558
max_lon <- -94.17598
mid_lat <- (min_lat + max_lat) / 2
mid_lon <- (min_lon + max_lon) / 2

# Box creator function
create_box <- function(coords, name) {
  poly <- st_polygon(list(matrix(coords, ncol = 2, byrow = TRUE)))
  st_sf(name = name, geometry = st_sfc(poly, crs = 4326))
}

# Main box
main_coords <- c(
  min_lon, min_lat, #bottom left
  min_lon, max_lat, #top left
  max_lon, max_lat, #top right
  max_lon, min_lat, #bottom right
  min_lon, min_lat #bottom left
)
main_box <- create_box(main_coords, "Overall Region")

# Southwest box
sw_coords <- c(
  -96.2, 28.9, #bottom left
  -96.2, 29.2, #top left
  -95.8, 29.2, #top right
  -95.8, 28.9, #bottom right
  -96.2, 28.9 #bottom left
)
sw_box <- create_box(sw_coords, "Box 1")

# Northeast box
ne_coords <- c(
  -94.8, 30, #bottom left
  -94.8, 30.3, #top left
  -94.4, 30.3, #top right
  -94.4, 30, #bottom right
  -94.8, 30 #bottom left
)
ne_box <- create_box(ne_coords, "Box 3")

# Top-left box
tl_coords <- c(
  -96.4, 30.2, #bottom left
  -96.4, 30.5, #top left
  -96, 30.5, #top right
  -96, 30.2, #bottom right
  -96.4, 30.2 #bottom left
)
tl_box <- create_box(tl_coords, "Box 2")

# Combine all into one sf object
all_boxes <- rbind(main_box, sw_box, ne_box, tl_box)

# Label positions (centroids of each box)
labels <- st_centroid(all_boxes)

# Plot in view mode
tmap_mode("view")
tm_shape(all_boxes) +
  tm_borders(lwd = 3, col = "blue") +
  tm_basemap("OpenStreetMap") +
  tm_fill("name",
        fill_alpha = 0.4,
        fill.scale = tm_scale(values = "brewer.set2")) + 
  tm_shape(labels) +
  tm_symbols(size = 0) +  # invisible symbol layer
  tm_text("name", size = 1.2, col = "black", options = opt_tm_text(point.label = TRUE))


```

#Now Look at Intersection

Now we want to look at the crop pixels within those three sub-boxes

```{r points within boxes}

#########################create the boxes

# Box 1 - bottom left
b1 <- matrix(c(
  -96.2, 28.9, #bottom left
  -96.2, 29.2, #top left
  -95.8, 29.2, #top right
  -95.8, 28.9, #bottom right
  -96.2, 28.9 #bottom left
), ncol=2, byrow=TRUE)

b1_poly <- st_polygon(list(b1)) |> st_sfc(crs = 4326)

# Box 2 - top left
b2 <- matrix(c(
  -96.4, 30.2, #bottom left
  -96.4, 30.5, #top left
  -96, 30.5, #top right
  -96, 30.2, #bottom right
  -96.4, 30.2 #bottom left
), ncol=2, byrow=TRUE)

b2_poly <- st_polygon(list(b2)) |> st_sfc(crs = 4326)

# Box 3 - top right
b3 <- matrix(c(
  -94.8, 30, #bottom left
  -94.8, 30.3, #top left
  -94.4, 30.3, #top right
  -94.4, 30, #bottom right
  -94.8, 30 #bottom left
), ncol=2, byrow=TRUE)

b3_poly <- st_polygon(list(b3)) |> st_sfc(crs = 4326)


####################create the points

#get the relevant info
pts_16 <- HourlyMeans16_0[, c("lat", "lon", "landuse")]
pts_17 <- HourlyMeans17_0[, c("lat", "lon", "landuse")]

##find the unique combos
unique_coords16 <- unique(pts_16[, c("lat", "lon")])
unique_coords17 <- unique(pts_17[, c("lat", "lon")])

#add label
unique_coords16$label <- paste0("Pt_", seq_len(nrow(unique_coords16)))
unique_coords17$label <- paste0("Pt_", seq_len(nrow(unique_coords17)))

#add them back to the other dfs too for later cross ref
HourlyMeans16_0 <- merge(HourlyMeans16_0, unique_coords16, by = c("lat", "lon"))
HourlyMeans16_1 <- merge(HourlyMeans16_1, unique_coords16, by = c("lat", "lon"))
HourlyMeans16_2 <- merge(HourlyMeans16_2, unique_coords16, by = c("lat", "lon"))

HourlyMeans17_0 <- merge(HourlyMeans17_0, unique_coords17, by = c("lat", "lon"))
HourlyMeans17_1 <- merge(HourlyMeans17_1, unique_coords17, by = c("lat", "lon"))
HourlyMeans17_2 <- merge(HourlyMeans17_2, unique_coords17, by = c("lat", "lon"))

#turn them into sf objects
pts_16_sf <- st_as_sf(unique_coords16, coords = c("lon", "lat"), crs = 4326)
pts_17_sf <- st_as_sf(unique_coords17, coords = c("lon", "lat"), crs = 4326)

#######################look at intersection
#check which are inside - box 1
inside_logical_16b1 <- st_within(pts_16_sf, b1_poly, sparse = FALSE)[, 1]
inside_logical_17b1 <- st_within(pts_17_sf, b1_poly, sparse = FALSE)[, 1]

pts_16_inside_b1 <- pts_16_sf[inside_logical_16b1, ]
pts_17_inside_b1 <- pts_17_sf[inside_logical_17b1, ]

#check which are inside - box 2
inside_logical_16b2 <- st_within(pts_16_sf, b2_poly, sparse = FALSE)[, 1]
inside_logical_17b2 <- st_within(pts_17_sf, b2_poly, sparse = FALSE)[, 1]

pts_16_inside_b2 <- pts_16_sf[inside_logical_16b2, ]
pts_17_inside_b2 <- pts_17_sf[inside_logical_17b2, ]

#check which are inside - box 3
inside_logical_16b3 <- st_within(pts_16_sf, b3_poly, sparse = FALSE)[, 1]
inside_logical_17b3 <- st_within(pts_17_sf, b3_poly, sparse = FALSE)[, 1]

pts_16_inside_b3 <- pts_16_sf[inside_logical_16b3, ]
pts_17_inside_b3 <- pts_17_sf[inside_logical_17b3, ]

```

##Now set up new datasets for analysis

```{r create new df for analysis}

###############create new df with box labels
#2016
HourlyMeans16_0wb <- HourlyMeans16_0 %>% 
  filter(landuse == "Crop") %>%
  mutate(region = case_when(
    label %in% pts_16_inside_b1$label ~ 1,
    label %in% pts_16_inside_b2$label ~ 2,
    label %in% pts_16_inside_b3$label ~ 3,
    TRUE ~ 0
  ))
HourlyMeans16_0wb$sim <- "sim1"

HourlyMeans16_1wb <- HourlyMeans16_1 %>% 
  filter(landuse == "Crop") %>%
  mutate(region = case_when(
    label %in% pts_16_inside_b1$label ~ 1,
    label %in% pts_16_inside_b2$label ~ 2,
    label %in% pts_16_inside_b3$label ~ 3,
    TRUE ~ 0
  ))
HourlyMeans16_1wb$sim <- "sim2"

HourlyMeans16_2wb <- HourlyMeans16_2 %>% 
  filter(landuse == "Crop") %>%
  mutate(region = case_when(
    label %in% pts_16_inside_b1$label ~ 1,
    label %in% pts_16_inside_b2$label ~ 2,
    label %in% pts_16_inside_b3$label ~ 3,
    TRUE ~ 0
  ))
HourlyMeans16_2wb$sim <- "sim3"

###2017
HourlyMeans17_0wb <- HourlyMeans17_0 %>% 
  filter(landuse == "Crop") %>%
  mutate(region = case_when(
    label %in% pts_17_inside_b1$label ~ 1,
    label %in% pts_17_inside_b2$label ~ 2,
    label %in% pts_17_inside_b3$label ~ 3,
    TRUE ~ 0
  ))
HourlyMeans17_0wb$sim <- "sim1"

HourlyMeans17_1wb <- HourlyMeans17_1 %>% 
  filter(landuse == "Crop") %>%
  mutate(region = case_when(
    label %in% pts_17_inside_b1$label ~ 1,
    label %in% pts_17_inside_b2$label ~ 2,
    label %in% pts_17_inside_b3$label ~ 3,
    TRUE ~ 0
  ))
HourlyMeans17_1wb$sim <- "sim2"

HourlyMeans17_2wb <- HourlyMeans17_2 %>% 
  filter(landuse == "Crop") %>%
  mutate(region = case_when(
    label %in% pts_17_inside_b1$label ~ 1,
    label %in% pts_17_inside_b2$label ~ 2,
    label %in% pts_17_inside_b3$label ~ 3,
    TRUE ~ 0
  ))
HourlyMeans17_2wb$sim <- "sim3"

```

```{r create dfs for combined plots}
#this is if you want to compare the dfs for the simulations versus the baseline for temperature, rh, slp, and wspd10

HourlyMeans16_wb <- bind_rows(HourlyMeans16_0wb, HourlyMeans16_1wb, HourlyMeans16_2wb)

HourlyMeans17_wb <- bind_rows(HourlyMeans17_0wb, HourlyMeans17_1wb, HourlyMeans17_2wb)

```



##Basic Plots

Create simple plots of temperature, relative humidity, and wind speed over time for crop pixels within the three boxes

```{r basic plots}
##Box 1 is bottom left
##Box 2 is top left
##Box 3 is top right

###################################Temperature
##2016 Basic Temperature Plot
ggplot(HourlyMeans16_wb[HourlyMeans16_wb$hour > 8 & HourlyMeans16_wb$region != 0,], aes(x=as.factor(hour), y=(T2-273.15), fill=as.factor(region))) + 
  geom_boxplot() + ggtitle("2016 Temperatures by Region") + 
  xlab("Hour") + ylab("Temperature (C)")  + 
   coord_cartesian(ylim = c(9, 25))

#2017
ggplot(HourlyMeans17_wb[HourlyMeans17_wb$hour > 8 & HourlyMeans17_wb$region != 0,], aes(x=as.factor(hour), y=(T2-273.15), fill=as.factor(region))) + 
  geom_boxplot() + ggtitle("2017 Temperatures by Region") + 
  xlab("Hour") + ylab("Temperature (C)")  + 
   coord_cartesian(ylim = c(9, 25))


###################################Humidity
##2016 Basic Plot
ggplot(HourlyMeans16_wb[HourlyMeans16_wb$hour > 8 & HourlyMeans16_wb$region != 0,], aes(x=as.factor(hour), y=(rh2), fill=as.factor(region))) + 
  geom_boxplot() + ggtitle("2016 Relative Humidity by Region") + 
  xlab("Hour") + ylab("Relative Humidity (%)")  + 
   coord_cartesian(ylim = c(0, 100))


##2017 Basic Plot
ggplot(HourlyMeans17_wb[HourlyMeans17_wb$hour > 8 & HourlyMeans17_wb$region != 0,], aes(x=as.factor(hour), y=(rh2), fill=as.factor(region))) + 
  geom_boxplot() + ggtitle("2017 Relative Humidity by Region") + 
  xlab("Hour") + ylab("Relative Humidity (%)")  + 
   coord_cartesian(ylim = c(0, 100))


###################################Wind
##2016 Basic Plot
ggplot(HourlyMeans16_wb[HourlyMeans16_wb$hour > 8 & HourlyMeans16_wb$region != 0,], aes(x=as.factor(hour), y=(wspd10), fill=as.factor(region))) + 
  geom_boxplot() + ggtitle("2016 Wind Speed by Region") + 
  xlab("Hour") + ylab("Wind Speed (m/s)")  + 
   coord_cartesian(ylim = c(0, 8))

##2017 Basic Plot
ggplot(HourlyMeans17_wb[HourlyMeans17_wb$hour > 8 & HourlyMeans17_wb$region != 0,], aes(x=as.factor(hour), y=(wspd10), fill=as.factor(region))) + 
  geom_boxplot() + ggtitle("2017 Wind Speed by Region") + 
  xlab("Hour") + ylab("Wind Speed (m/s)")  + 
   coord_cartesian(ylim = c(0, 8))

###################################Low level clouds
##2016 Basic Plot
ggplot(HourlyMeans16_wb[HourlyMeans16_wb$hour > 8 & HourlyMeans16_wb$region != 0,], aes(x=as.factor(hour), y=(llc), fill=as.factor(region))) + 
  geom_boxplot() + ggtitle("2016 Low-Level Clouds by Region") + 
  xlab("Hour") + ylab("Low level cloud cover %")  + 
   coord_cartesian(ylim = c(0, 0.4))

##2017 Basic Plot
ggplot(HourlyMeans17_wb[HourlyMeans17_wb$hour > 8 & HourlyMeans17_wb$region != 0,], aes(x=as.factor(hour), y=(llc), fill=as.factor(region))) + 
  geom_boxplot() + ggtitle("2017 Low-Level Clouds by Region") + 
  xlab("Hour") + ylab("Low level cloud cover %")  + 
   coord_cartesian(ylim = c(0, 0.4))


```

#Next step

Should probably make this more sophisticated by handling the simulations differently too, right now they all are just thrown in together 

```{r reformat to compare across years}

#for final storage matrix
colnms <- c("Hour", "Box", "Temp16", "Temp17", "TempDiff",
            "RelHum16", "RelHum17", "HumDif17m16",
            "WindSpeed10m16", "WindSpeed10m17", "WSDiff", 
            "LowLvlCl16", "LowLvlCl17", "LLCDiff")

#Create df
BoxMeans <- as.data.frame(matrix(0, nrow = 24*3, ncol=length(colnms)))
names(BoxMeans) <- colnms

iter <- 1

BM16 <- HourlyMeans16_wb %>% 
  summarize(MeanTemp16 = mean(T2),
            MeanRH16 = mean(rh2),
            MeanWspd16 = mean(wspd10),
            MeanLLC16 = mean(llc), .by=c("hour", "region", "sim"))

BM17 <- HourlyMeans17_wb %>% 
  summarize(MeanTemp17 = mean(T2),
            MeanRH17 = mean(rh2),
            MeanWspd17 = mean(wspd10),
            MeanLLC17 = mean(llc), .by=c("hour", "region", "sim"))

BoxMeans <- BM17 %>% left_join(BM16, by = c('hour', 'region', 'sim')) %>%
  mutate(TempDiff = MeanTemp17 - MeanTemp16,
         RHDiff = MeanRH17 - MeanRH16,
         WspdDiff = MeanWspd17 - MeanWspd16,
         LLCDiff = MeanLLC17 - MeanLLC16)


```


##Difference Visuals
```{r diff visuals temp}

###difference plots
ggplot(BoxMeans, aes(x=as.factor(hour), y=(TempDiff), group=region)) + 
  geom_point(aes(color=as.factor(region))) + ggtitle("2017 Minus 2016 Temperatures by Region") + 
  xlab("hour") + ylab("Temperature Difference (C)")  + geom_hline(yintercept=0, linetype="dashed") + 
   coord_cartesian(ylim = c(-1, 2)) + theme_minimal() 


####Simple figures to make sure the individual averages over time are reasonable

ggplot(BoxMeans, aes(x=as.factor(hour), y=(MeanTemp16-273.15), group=region)) + 
  geom_point(aes(color=as.factor(region))) + ggtitle("2016 Temperatures by Region") + 
  xlab("hour") + ylab("Temperature (C)") + 
   coord_cartesian(ylim = c(4, 24)) + theme_minimal() 

ggplot(BoxMeans, aes(x=as.factor(hour), y=(MeanTemp17-273.15), group=region)) + 
  geom_point(aes(color=as.factor(region))) + ggtitle("2017 Temperatures by Region") + 
  xlab("hour") + ylab("Temperature (C)") + 
   coord_cartesian(ylim = c(4, 24)) + theme_minimal() 

```

Now for relative humidity

```{r diff visuals RH}

##difference plots
ggplot(BoxMeans, aes(x=as.factor(hour), y=(RHDiff), group=region)) + 
  geom_point(aes(color=as.factor(region))) + ggtitle("2017 Minus 2016 Relative Humidity by Region") + 
  xlab("hour") + ylab("RH difference Difference (%)")  + geom_hline(yintercept=0, linetype="dashed") + 
   coord_cartesian(ylim = c(-15, 5))+ theme_minimal() 



###individual plots for each year to make sure they are reasonable
ggplot(BoxMeans, aes(x=as.factor(hour), y=(MeanRH16), group=region)) + 
  geom_point(aes(color=as.factor(region))) + ggtitle("2016 Relative Humidity by Region") + 
  xlab("hour") + ylab("RH (%)") + geom_hline(yintercept=0, linetype="dashed")  + 
   coord_cartesian(ylim = c(0, 100))+ theme_minimal() 

ggplot(BoxMeans, aes(x=as.factor(hour), y=(MeanRH17), group=region)) + 
  geom_point(aes(color=as.factor(region))) + ggtitle("2017 Relative Humidity by Region") + 
  xlab("hour") + ylab("RH (%)") + geom_hline(yintercept=0, linetype="dashed")  + 
   coord_cartesian(ylim = c(0, 100))+ theme_minimal() 

```

Now for wind speed

```{r diff visuals wind speed}
#difference plots
ggplot(BoxMeans, aes(x=as.factor(hour), y=(WspdDiff), group=region)) + 
  geom_point(aes(color=as.factor(region))) + ggtitle("2017 Minus 2016 Wind Speed by Region") + 
  xlab("hour") + ylab("Wind Speed difference Difference (m/s)")  + geom_hline(yintercept=0, linetype="dashed") + 
   coord_cartesian(ylim = c(-1, 5))+ theme_minimal() 


###individual plots for each year to make sure they make sense
ggplot(BoxMeans, aes(x=as.factor(hour), y=(MeanWspd16), group=region)) + 
  geom_point(aes(color=as.factor(region))) + ggtitle("2016 Wind Speed by Region") + 
  xlab("hour") + ylab("Wind speed (m/s)") + geom_hline(yintercept=0, linetype="dashed") + 
   coord_cartesian(ylim = c(0, 18))+ theme_minimal() 

ggplot(BoxMeans, aes(x=as.factor(hour), y=(MeanWspd17), group=region)) + 
  geom_point(aes(color=as.factor(region))) + ggtitle("2017 Wind Speed by Region") + 
  xlab("hour") + ylab("Wind speed (m/s)") + geom_hline(yintercept=0, linetype="dashed") + 
   coord_cartesian(ylim = c(0, 18))+ theme_minimal() 


```


## Low-level Clouds
```{r diff in llc across time}
ggplot(BoxMeans, aes(x=as.factor(hour), y=(LLCDiff), group=region)) + 
  geom_point(aes(color=as.factor(region))) + ggtitle("2017 Minus 2016 low-level clouds by Region") + 
  xlab("hour") + ylab("Low Level Clouds Difference")  + geom_hline(yintercept=0, linetype="dashed") + theme_minimal() 

#Plots for each year for reference
ggplot(BoxMeans, aes(x=as.factor(hour), y=(MeanLLC16), group=region)) + 
  geom_point(aes(color=as.factor(region))) + ggtitle("Low-level Clouds by Region 2016") + 
  xlab("hour") + ylab("% Low-level clouds") + geom_hline(yintercept=0, linetype="dashed") + 
   coord_cartesian(ylim = c(0, 0.1))+ theme_minimal() 

ggplot(BoxMeans, aes(x=as.factor(hour), y=(MeanLLC17), group=region)) + 
  geom_point(aes(color=as.factor(region))) + ggtitle("Low-level Clouds by Region 2017") + 
  xlab("hour") + ylab("% Low-level clouds") + geom_hline(yintercept=0, linetype="dashed") + 
   coord_cartesian(ylim = c(0, 0.1))+ theme_minimal() 

```