---
title: "MHW Spatial EDA"
author: "Renata Poulton Kamakura"
date: "2025-05-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(sf)
library(tmap)
library(dplyr)
```

## Input data

Pulling in both the hourly means and the distance to ocean csv files with some simplification of the land use/land cover types to make it more manageable

```{r read in data}

#basic wrfoutput data
HourlyMeans16_0 <- read.csv("../../03ProcessedData/HourlyMeans_2016_0.csv")
HourlyMeans16_1 <- read.csv("../../03ProcessedData/HourlyMeans_2016_1.csv")
HourlyMeans16_2 <- read.csv("../../03ProcessedData/HourlyMeans_2016_2.csv")

HourlyMeans17_0 <- read.csv("../../03ProcessedData/HourlyMeans_2017_0.csv")
HourlyMeans17_1 <- read.csv("../../03ProcessedData/HourlyMeans_2017_1.csv")
HourlyMeans17_2 <- read.csv("../../03ProcessedData/HourlyMeans_2017_1.csv")

#to simplify the land use types
LU_names = c('Evergreen Needleleaf Forest', 'Evergreen Broadleaf Forest', 
            'Deciduous Needleleaf Forest', 'Deciduous Broadleaf Forest',
            'Mixed Forests', 'Closed Shrublands',  'Open Shrublands', 
            'Woody Savannas', 'Savannas', 'Grasslands', 
            'Permanent wetlands', 'Croplands', 'Urban and Built-Up', 
            'cropland/natural vegetation mosaic', 
            'Snow and Ice', 'Barren or Sparsely Vegetated', 'Water')

NewLUs <- c('Forest', 'Forest', 'Forest', 'Forest', 'Forest',
            'Shrub/Savanna', 'Shrub/Savanna', 'Shrub/Savanna', 'Shrub/Savanna',
            "Shrub/Savanna", "Wetlands", "Crop", "Urban", "Crop", "Snow", 
            "Barren", "Water")
NewLU_nums <- c(1,1,1,1,1,2,2,2,2,2,3,4,5,4,6,7,8)

LU_dict <- as.data.frame(matrix(0, nrow=17, ncol=4))
names(LU_dict) <- c("Original_Name", "Original_Num", "New_Name", "New_Num")

LU_dict$Original_Name <- LU_names
LU_dict$Original_Num <- 1:17
LU_dict$New_Name <- NewLUs
LU_dict$New_Num <- NewLU_nums

#distance from ocean pixels
Dists16 <- read.csv("../../03ProcessedData/Dist2Ocean/Dist2Ocean_2016_0.csv")
Dists17 <- read.csv("../../03ProcessedData/Dist2Ocean/Dist2Ocean_2017_0.csv")

##combine
HourlyMeans16_0$Dist2Ocean <- Dists16$Dist2Ocean[Dists16$lat == HourlyMeans16_0$lat & Dists16$long == HourlyMeans16_0$lon]
HourlyMeans16_1$Dist2Ocean <- Dists16$Dist2Ocean[Dists16$lat == HourlyMeans16_1$lat & Dists16$long == HourlyMeans16_1$lon]
HourlyMeans16_2$Dist2Ocean <- Dists16$Dist2Ocean[Dists16$lat == HourlyMeans16_2$lat & Dists16$long == HourlyMeans16_2$lon]

HourlyMeans17_0$Dist2Ocean <- Dists17$Dist2Ocean[Dists16$lat == HourlyMeans17_0$lat & Dists17$long == HourlyMeans17_0$lon]
HourlyMeans17_1$Dist2Ocean <- Dists17$Dist2Ocean[Dists16$lat == HourlyMeans17_1$lat & Dists17$long == HourlyMeans17_1$lon]
HourlyMeans17_2$Dist2Ocean <- Dists17$Dist2Ocean[Dists16$lat == HourlyMeans17_2$lat & Dists17$long == HourlyMeans17_2$lon]


#########change the hours to local time zone from UTC
HourDict <- as.data.frame(matrix(0, nrow=24, ncol=2))
names(HourDict) <- c("Current", "New")
HourDict$Current <- 0:23
HourDict$New <- c(18, 19, 20, 21, 22, 23, 0, 1, 2, 
                  3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 
                  13, 14, 15, 16, 17)

#convert the hours to CST
HourlyMeans16_0$hour <- HourDict$New[HourlyMeans16_0$hour + 1] 
HourlyMeans16_1$hour <- HourDict$New[HourlyMeans16_1$hour + 1] 
HourlyMeans16_2$hour <- HourDict$New[HourlyMeans16_2$hour + 1] 

HourlyMeans17_0$hour <- HourDict$New[HourlyMeans17_0$hour + 1] 
HourlyMeans17_1$hour <- HourDict$New[HourlyMeans17_1$hour + 1] 
HourlyMeans17_2$hour <- HourDict$New[HourlyMeans17_2$hour + 1] 

#########modify land use
HourlyMeans16_0$landuse <- LU_dict$New_Name[HourlyMeans16_0$landuse]
HourlyMeans16_1$landuse <- LU_dict$New_Name[HourlyMeans16_1$landuse]
HourlyMeans16_2$landuse <- LU_dict$New_Name[HourlyMeans16_2$landuse]

HourlyMeans17_0$landuse <- LU_dict$New_Name[HourlyMeans17_0$landuse]
HourlyMeans17_1$landuse <- LU_dict$New_Name[HourlyMeans17_1$landuse]
HourlyMeans17_2$landuse <- LU_dict$New_Name[HourlyMeans17_2$landuse]

```


#basic visuals of your regions

As background reference to help orient the user to what they will be looking at

```{r basic vis}

# Coordinates
min_lat <- 28.70586
max_lat <- 30.68516
min_lon <- -96.50558
max_lon <- -94.17598
mid_lat <- (min_lat + max_lat) / 2
mid_lon <- (min_lon + max_lon) / 2

# Box creator function
create_box <- function(coords, name) {
  poly <- st_polygon(list(matrix(coords, ncol = 2, byrow = TRUE)))
  st_sf(name = name, geometry = st_sfc(poly, crs = 4326))
}

# Main box
main_coords <- c(
  min_lon, min_lat, #bottom left
  min_lon, max_lat, #top left
  max_lon, max_lat, #top right
  max_lon, min_lat, #bottom right
  min_lon, min_lat #bottom left
)
main_box <- create_box(main_coords, "Overall Region")

# Southwest box
sw_coords <- c(
  -96.2, 28.9, #bottom left
  -96.2, 29.2, #top left
  -95.8, 29.2, #top right
  -95.8, 28.9, #bottom right
  -96.2, 28.9 #bottom left
)
sw_box <- create_box(sw_coords, "Box 1")

# Northeast box
ne_coords <- c(
  -94.8, 30, #bottom left
  -94.8, 30.3, #top left
  -94.4, 30.3, #top right
  -94.4, 30, #bottom right
  -94.8, 30 #bottom left
)
ne_box <- create_box(ne_coords, "Box 3")

# Top-left box
tl_coords <- c(
  -96.4, 30.2, #bottom left
  -96.4, 30.5, #top left
  -96, 30.5, #top right
  -96, 30.2, #bottom right
  -96.4, 30.2 #bottom left
)
tl_box <- create_box(tl_coords, "Box 2")

# Combine all into one sf object
all_boxes <- rbind(main_box, sw_box, ne_box, tl_box)

# Label positions (centroids of each box)
labels <- st_centroid(all_boxes)

# Plot in view mode
tmap_mode("view")
tm_shape(all_boxes) +
  tm_borders(lwd = 3, col = "blue") +
  tm_basemap("OpenStreetMap") +
  tm_fill("name",
        fill_alpha = 0.4,
        fill.scale = tm_scale(values = "brewer.set2")) + 
  tm_shape(labels) +
  tm_symbols(size = 0) +  # invisible symbol layer
  tm_text("name", size = 1.2, col = "black", options = opt_tm_text(point.label = TRUE))


```

#Now Look at Intersection

Now we want to look at the crop pixels within those three sub-boxes

```{r points within boxes}

#########################create the boxes

# Box 1 - bottom left
b1 <- matrix(c(
  -96.2, 28.9, #bottom left
  -96.2, 29.2, #top left
  -95.8, 29.2, #top right
  -95.8, 28.9, #bottom right
  -96.2, 28.9 #bottom left
), ncol=2, byrow=TRUE)

b1_poly <- st_polygon(list(b1)) |> st_sfc(crs = 4326)

# Box 2 - top left
b2 <- matrix(c(
  -96.4, 30.2, #bottom left
  -96.4, 30.5, #top left
  -96, 30.5, #top right
  -96, 30.2, #bottom right
  -96.4, 30.2 #bottom left
), ncol=2, byrow=TRUE)

b2_poly <- st_polygon(list(b2)) |> st_sfc(crs = 4326)

# Box 3 - top right
b3 <- matrix(c(
  -94.8, 30, #bottom left
  -94.8, 30.3, #top left
  -94.4, 30.3, #top right
  -94.4, 30, #bottom right
  -94.8, 30 #bottom left
), ncol=2, byrow=TRUE)

b3_poly <- st_polygon(list(b3)) |> st_sfc(crs = 4326)


####################create the points

#get the relevant info
pts_16 <- HourlyMeans16_0[, c("lat", "lon", "landuse")]
pts_17 <- HourlyMeans17_0[, c("lat", "lon", "landuse")]

##find the unique combos
unique_coords16 <- unique(pts_16[, c("lat", "lon")])
unique_coords17 <- unique(pts_17[, c("lat", "lon")])

#add label
unique_coords16$label <- paste0("Pt_", seq_len(nrow(unique_coords16)))
unique_coords17$label <- paste0("Pt_", seq_len(nrow(unique_coords17)))

#add them back to the other dfs too for later cross ref
HourlyMeans16_0 <- merge(HourlyMeans16_0, unique_coords16, by = c("lat", "lon"))
HourlyMeans16_1 <- merge(HourlyMeans16_1, unique_coords16, by = c("lat", "lon"))
HourlyMeans16_2 <- merge(HourlyMeans16_2, unique_coords16, by = c("lat", "lon"))

HourlyMeans17_0 <- merge(HourlyMeans17_0, unique_coords17, by = c("lat", "lon"))
HourlyMeans17_1 <- merge(HourlyMeans17_1, unique_coords17, by = c("lat", "lon"))
HourlyMeans17_2 <- merge(HourlyMeans17_2, unique_coords17, by = c("lat", "lon"))

#turn them into sf objects
pts_16_sf <- st_as_sf(unique_coords16, coords = c("lon", "lat"), crs = 4326)
pts_17_sf <- st_as_sf(unique_coords17, coords = c("lon", "lat"), crs = 4326)

#######################look at intersection
#check which are inside - box 1
inside_logical_16b1 <- st_within(pts_16_sf, b1_poly, sparse = FALSE)[, 1]
inside_logical_17b1 <- st_within(pts_17_sf, b1_poly, sparse = FALSE)[, 1]

pts_16_inside_b1 <- pts_16_sf[inside_logical_16b1, ]
pts_17_inside_b1 <- pts_17_sf[inside_logical_17b1, ]

#check which are inside - box 2
inside_logical_16b2 <- st_within(pts_16_sf, b2_poly, sparse = FALSE)[, 1]
inside_logical_17b2 <- st_within(pts_17_sf, b2_poly, sparse = FALSE)[, 1]

pts_16_inside_b2 <- pts_16_sf[inside_logical_16b2, ]
pts_17_inside_b2 <- pts_17_sf[inside_logical_17b2, ]

#check which are inside - box 3
inside_logical_16b3 <- st_within(pts_16_sf, b3_poly, sparse = FALSE)[, 1]
inside_logical_17b3 <- st_within(pts_17_sf, b3_poly, sparse = FALSE)[, 1]

pts_16_inside_b3 <- pts_16_sf[inside_logical_16b3, ]
pts_17_inside_b3 <- pts_17_sf[inside_logical_17b3, ]

```

##Now set up new datasets for analysis

```{r create new df for analysis}

###############create new df with box labels
#2016
HourlyMeans16_0wb <- HourlyMeans16_0 %>% 
  filter(landuse == "Crop") %>%
  mutate(region = case_when(
    label %in% pts_16_inside_b1$label ~ 1,
    label %in% pts_16_inside_b2$label ~ 2,
    label %in% pts_16_inside_b3$label ~ 3,
    TRUE ~ 0
  ))
HourlyMeans16_0wb$sim <- "sim1"

HourlyMeans16_1wb <- HourlyMeans16_1 %>% 
  filter(landuse == "Crop") %>%
  mutate(region = case_when(
    label %in% pts_16_inside_b1$label ~ 1,
    label %in% pts_16_inside_b2$label ~ 2,
    label %in% pts_16_inside_b3$label ~ 3,
    TRUE ~ 0
  ))
HourlyMeans16_1wb$sim <- "sim2"

HourlyMeans16_2wb <- HourlyMeans16_2 %>% 
  filter(landuse == "Crop") %>%
  mutate(region = case_when(
    label %in% pts_16_inside_b1$label ~ 1,
    label %in% pts_16_inside_b2$label ~ 2,
    label %in% pts_16_inside_b3$label ~ 3,
    TRUE ~ 0
  ))
HourlyMeans16_2wb$sim <- "sim3"

###2017
HourlyMeans17_0wb <- HourlyMeans17_0 %>% 
  filter(landuse == "Crop") %>%
  mutate(region = case_when(
    label %in% pts_17_inside_b1$label ~ 1,
    label %in% pts_17_inside_b2$label ~ 2,
    label %in% pts_17_inside_b3$label ~ 3,
    TRUE ~ 0
  ))
HourlyMeans17_0wb$sim <- "sim1"

HourlyMeans17_1wb <- HourlyMeans17_1 %>% 
  filter(landuse == "Crop") %>%
  mutate(region = case_when(
    label %in% pts_17_inside_b1$label ~ 1,
    label %in% pts_17_inside_b2$label ~ 2,
    label %in% pts_17_inside_b3$label ~ 3,
    TRUE ~ 0
  ))
HourlyMeans17_1wb$sim <- "sim2"

HourlyMeans17_2wb <- HourlyMeans17_2 %>% 
  filter(landuse == "Crop") %>%
  mutate(region = case_when(
    label %in% pts_17_inside_b1$label ~ 1,
    label %in% pts_17_inside_b2$label ~ 2,
    label %in% pts_17_inside_b3$label ~ 3,
    TRUE ~ 0
  ))
HourlyMeans17_2wb$sim <- "sim3"

```

```{r create dfs for combined plots}
#this is if you want to compare the dfs for the simulations versus the baseline for temperature, rh, slp, and wspd10

HourlyMeans16_wb <- bind_rows(HourlyMeans16_0wb, HourlyMeans16_1wb, HourlyMeans16_2wb)

HourlyMeans17_wb <- bind_rows(HourlyMeans17_0wb, HourlyMeans17_1wb, HourlyMeans17_2wb)

```



##Basic Plots

Create simple plots of temperature, relative humidity, and wind speed over time for crop pixels within the three boxes

```{r basic plots}
##Box 1 is bottom left
##Box 2 is top left
##Box 3 is top right

###################################Temperature
##2016 Basic Temperature Plot
ggplot(HourlyMeans16_wb[HourlyMeans16_wb$hour > 8 & HourlyMeans16_wb$region != 0,], aes(x=as.factor(hour), y=(T2-273.15), fill=as.factor(region))) + 
  geom_boxplot() + ggtitle("2016 Temperatures by Region") + 
  xlab("Hour") + ylab("Temperature (C)")  + 
   coord_cartesian(ylim = c(9, 25))

#2017
ggplot(HourlyMeans17_wb[HourlyMeans17_wb$hour > 8 & HourlyMeans17_wb$region != 0,], aes(x=as.factor(hour), y=(T2-273.15), fill=as.factor(region))) + 
  geom_boxplot() + ggtitle("2017 Temperatures by Region") + 
  xlab("Hour") + ylab("Temperature (C)")  + 
   coord_cartesian(ylim = c(9, 25))


###################################Humidity
##2016 Basic Plot
ggplot(HourlyMeans16_wb[HourlyMeans16_wb$hour > 8 & HourlyMeans16_wb$region != 0,], aes(x=as.factor(hour), y=(rh2), fill=as.factor(region))) + 
  geom_boxplot() + ggtitle("2016 Relative Humidity by Region") + 
  xlab("Hour") + ylab("Relative Humidity (%)")  + 
   coord_cartesian(ylim = c(0, 100))


##2017 Basic Plot
ggplot(HourlyMeans17_wb[HourlyMeans17_wb$hour > 8 & HourlyMeans17_wb$region != 0,], aes(x=as.factor(hour), y=(rh2), fill=as.factor(region))) + 
  geom_boxplot() + ggtitle("2017 Relative Humidity by Region") + 
  xlab("Hour") + ylab("Relative Humidity (%)")  + 
   coord_cartesian(ylim = c(0, 100))


###################################Wind
##2016 Basic Plot
ggplot(HourlyMeans16_wb[HourlyMeans16_wb$hour > 8 & HourlyMeans16_wb$region != 0,], aes(x=as.factor(hour), y=(wspd10), fill=as.factor(region))) + 
  geom_boxplot() + ggtitle("2016 Wind Speed by Region") + 
  xlab("Hour") + ylab("Wind Speed (m/s)")  + 
   coord_cartesian(ylim = c(0, 8))

##2017 Basic Plot
ggplot(HourlyMeans17_wb[HourlyMeans17_wb$hour > 8 & HourlyMeans17_wb$region != 0,], aes(x=as.factor(hour), y=(wspd10), fill=as.factor(region))) + 
  geom_boxplot() + ggtitle("2017 Wind Speed by Region") + 
  xlab("Hour") + ylab("Wind Speed (m/s)")  + 
   coord_cartesian(ylim = c(0, 8))



```

#Next step
Now I would look at the three boxes comparing MHW to non-MHW both ways

I suspect you will see a bigger difference between the two 2017 runs for box 2, where the wind probably has to go through the city and is disrupted, compared to what you see in the boxes on either side of the city 

Note, all of these are labelled as crop-ish land use

```{r reformat to compare across years}

#for final storage matrix
colnms <- c("Hour", "Box", "Temp16", "Temp17", "Temp17s", "TempDiff17m16", "TempDiff17m17s", 
            "RelHum16", "RelHum17", "RelHum17s", "HumDiff17m16", "HumDiff17m17s",
            "WindSpeed10m16", "WindSpeed10m17", "WindSpeed10m17s", "WSDiff17m16", "WSDiff17m17s")

#Create df
BoxMeans <- as.data.frame(matrix(0, nrow = 24*3, ncol=length(colnms)))
names(BoxMeans) <- colnms

iter <- 1

for (h in c(0:23)){ #hours
  
  for (b in c(1:3)){ #boxes
    
    BoxMeans$Hour[iter] <- h
    BoxMeans$Box[iter] <- b
    
    #############Temperature
    BoxMeans$Temp16[iter] <- mean(HourlyMeans16_wBoxes$X2mTemp[HourlyMeans16_wBoxes$hour == h & HourlyMeans16_wBoxes$Box == b])
    BoxMeans$Temp17[iter] <- mean(HourlyMeans17_wBoxes$X2mTemp[HourlyMeans17_wBoxes$hour == h & HourlyMeans17_wBoxes$Box == b])
    BoxMeans$Temp17s[iter] <- mean(HourlyMeans17s_wBoxes$X2mTemp[HourlyMeans17s_wBoxes$hour == h & HourlyMeans17s_wBoxes$Box == b])
    
    #Calculate the difference
    BoxMeans$TempDiff17m16[iter] <- BoxMeans$Temp17[iter] - BoxMeans$Temp16[iter]
    BoxMeans$TempDiff17m17s[iter] <- BoxMeans$Temp17[iter] - BoxMeans$Temp17s[iter]
    
    ###############Humidity
    BoxMeans$RelHum16[iter] <- mean(HourlyMeans16_wBoxes$RelHumid[HourlyMeans16_wBoxes$hour == h & HourlyMeans16_wBoxes$Box == b])
    BoxMeans$RelHum17[iter] <- mean(HourlyMeans17_wBoxes$RelHumid[HourlyMeans17_wBoxes$hour == h & HourlyMeans17_wBoxes$Box == b])
    BoxMeans$RelHum17s[iter] <- mean(HourlyMeans17s_wBoxes$RelHumid[HourlyMeans17s_wBoxes$hour == h & HourlyMeans17s_wBoxes$Box == b])
    
    #Calculate the difference
    BoxMeans$HumDiff17m16[iter] <- BoxMeans$RelHum17[iter] - BoxMeans$RelHum16[iter]
    BoxMeans$HumDiff17m17s[iter] <- BoxMeans$RelHum17[iter] - BoxMeans$RelHum17s[iter]
    
    
    #################Wind Speed
    BoxMeans$WindSpeed10m16[iter] <- mean(HourlyMeans16_wBoxes$WindSpeed_10m[HourlyMeans16_wBoxes$hour == h & HourlyMeans16_wBoxes$Box == b])
    BoxMeans$WindSpeed10m17[iter] <- mean(HourlyMeans17_wBoxes$WindSpeed_10m[HourlyMeans17_wBoxes$hour == h & HourlyMeans17_wBoxes$Box == b])
    BoxMeans$WindSpeed10m17s[iter] <- mean(HourlyMeans17s_wBoxes$WindSpeed_10m[HourlyMeans17s_wBoxes$hour == h & HourlyMeans17s_wBoxes$Box == b])
    
    #Calculate the difference
    BoxMeans$WSDiff17m16[iter] <- BoxMeans$WindSpeed10m17[iter] - BoxMeans$WindSpeed10m16[iter]
    BoxMeans$WSDiff17m17s[iter] <- BoxMeans$WindSpeed10m17[iter] - BoxMeans$WindSpeed10m17s[iter]
    
    #increase the iterator
    iter <- iter + 1
  }
  
}


```