---
title: "WRF Output EDA"
author: "Renata Poulton Kamakura"
date: "2025-02-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(spBayes) #spatial random effects
library(geoR) #variogram functions
```

##Read in data

This is currently an incomplete run

```{r read data}
#Get the hourly averages for each year
#Likely will need to update the files here once re-run with all the data and spin-up
HourlyMeans16 <- read.csv("../03ProcessedData/HourlyMeans_2016_021025.csv")
HourlyMeans17 <- read.csv("../03ProcessedData/HourlyMeans_2017_021025.csv")

#More for reference
LU_names = c('Evergreen Needleleaf Forest', 'Evergreen Broadleaf Forest', 
            'Deciduous Needleleaf Forest', 'Deciduous Broadleaf Forest',
            'Mixed Forests', 'Closed Shrublands',  'Open Shrublands', 
            'Woody Savannas', 'Savannas', 'Grasslands', 
            'Permanent wetlands', 'Croplands', 'Urban and Built-Up', 
            'cropland/natural vegetation mosaic', 
            'Snow and Ice', 'Barren or Sparsely Vegetated', 'Water')

NewLUs <- c('Forest', 'Forest', 'Forest', 'Forest', 'Forest',
            'Shrub/Savanna', 'Shrub/Savanna', 'Shrub/Savanna', 'Shrub/Savanna',
            "Shrub/Savanna", "Wetlands", "Crop", "Urban", "Crop", "Snow", 
            "Barren", "Water")
NewLU_nums <- c(1,1,1,1,1,2,2,2,2,2,3,4,5,4,6,7,8)

LU_dict <- as.data.frame(matrix(0, nrow=17, ncol=4))
names(LU_dict) <- c("Original_Name", "Original_Num", "New_Name", "New_Num")

LU_dict$Original_Name <- LU_names
LU_dict$Original_Num <- 1:17
LU_dict$New_Name <- NewLUs
LU_dict$New_Num <- NewLU_nums

```

##Basic EDA

Ideally you want a linear mixed effects model. You can either run it with spatial random effects or use grid cell as a random intercept. My guess is that you should try running it with a spatial random effect but should try to look at the data first 

From looking at this, I suspect the hourly means are not working like they should, these seem too homogenous...

```{r temp EDA}
#A little easier to mess around with
aftHours1 <- c(12,13,14,15,16) #earlier as you get the start of the sea breeze
aftHours2 <- c(17,18,19,20,21) #as the land starts to cool down 

AfternoonMeans16_1 <- HourlyMeans16[HourlyMeans16$hour %in% aftHours1,]
AfternoonMeans17_1 <- HourlyMeans17[HourlyMeans17$hour %in% aftHours1,]

AfternoonMeans16_2 <- HourlyMeans16[HourlyMeans16$hour %in% aftHours2,]
AfternoonMeans17_2 <- HourlyMeans17[HourlyMeans17$hour %in% aftHours2,]

##########################temp and lat
ggplot(AfternoonMeans16_1, aes(x=lat, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Early Afternoon Temperatures with Latitude") + 
  xlab("Latitude") + ylab("Temperature (C)")

ggplot(AfternoonMeans16_2, aes(x=lat, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Late Afternoon Temperatures with Latitude") + 
  xlab("Latitude") + ylab("Temperature (C)")

ggplot(AfternoonMeans17_1, aes(x=lat, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Early Afternoon Temperatures with Latitude") + 
  xlab("Latitude") + ylab("Temperature (C)")

ggplot(AfternoonMeans17_2, aes(x=lat, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Late Afternoon Temperatures with Latitude") + 
  xlab("Latitude") + ylab("Temperature (C)")

##########################temp and long
ggplot(AfternoonMeans16_1, aes(x=long, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Early Afternoon Temperatures with Longitude") + 
  xlab("Longitude") + ylab("Temperature (C)")

ggplot(AfternoonMeans16_2, aes(x=long, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Late Afternoon Temperatures with Longitude") + 
  xlab("Longitude") + ylab("Temperature (C)")

ggplot(AfternoonMeans17_1, aes(x=long, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Early Afternoon Temperatures with Longitude") + 
  xlab("Longitude") + ylab("Temperature (C)")

ggplot(AfternoonMeans17_2, aes(x=long, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Late Afternoon Temperatures with Longitude") + 
  xlab("Longitude") + ylab("Temperature (C)")

############################temp and land use + hour
AfternoonMeans16_modLU_1 <- AfternoonMeans16_1
AfternoonMeans16_modLU_1$LandUse <- LU_dict$New_Name[AfternoonMeans16_1$LandUse]
AfternoonMeans16_modLU_2 <- AfternoonMeans16_2
AfternoonMeans16_modLU_2$LandUse<- LU_dict$New_Name[AfternoonMeans16_2$LandUse]

AfternoonMeans17_modLU_1 <- AfternoonMeans17_1
AfternoonMeans17_modLU_1$LandUse <- LU_dict$New_Name[AfternoonMeans17_1$LandUse]
AfternoonMeans17_modLU_2 <- AfternoonMeans17_2
AfternoonMeans17_modLU_2$LandUse <- LU_dict$New_Name[AfternoonMeans17_2$LandUse] 

##ideally should group land use somehow
ggplot(AfternoonMeans16_modLU_1, aes(x=as.factor(hour), y=(X2mTemp-273.15), fill=LandUse)) + 
  geom_boxplot() + ggtitle("2016 Early Afternoon Temperatures by Land Use") + 
  xlab("Hour") + ylab("Temperature (C)")

ggplot(AfternoonMeans16_modLU_2, aes(x=as.factor(hour), y=(X2mTemp-273.15), fill=LandUse)) + 
  geom_boxplot() + ggtitle("2016 Late Afternoon Temperatures by Land Use") + 
  xlab("Hour") + ylab("Temperature (C)")


ggplot(AfternoonMeans17_modLU_1, aes(x=as.factor(hour), y=(X2mTemp-273.15), fill=LandUse)) + 
  geom_boxplot() + ggtitle("2017 Early Afternoon Temperatures by Land Use") + 
  xlab("Hour") + ylab("Temperature (C)")

ggplot(AfternoonMeans17_modLU_2, aes(x=as.factor(hour), y=(X2mTemp-273.15), fill=LandUse)) + 
  geom_boxplot() + ggtitle("2017 Late Afternoon Temperatures by Land Use") + 
  xlab("Hour") + ylab("Temperature (C)")

############################temp and distance to ocean

ggplot(AfternoonMeans16_1, aes(x=Dist2Ocean, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Early Afternoon Temperatures with Distance to Ocean") + 
  xlab("Distance to Ocean (m)") + ylab("Temperature (C)")

ggplot(AfternoonMeans16_2, aes(x=Dist2Ocean, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Late Afternoon Temperatures with Distance to Ocean") + 
  xlab("Distance to Ocean (m)") + ylab("Temperature (C)")

ggplot(AfternoonMeans17_1, aes(x=Dist2Ocean, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Early Afternoon Temperatures with Distance to Ocean") + 
  xlab("Distance to Ocean (m)") + ylab("Temperature (C)")

ggplot(AfternoonMeans17_2, aes(x=Dist2Ocean, y=(X2mTemp-273.15), color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Late Afternoon Temperatures with Distance to Ocean") + 
  xlab("Distance to Ocean (m)") + ylab("Temperature (C)")
```


```{r humidity eda}

####################Relative Humidity


#rel humidity and lat
ggplot(AfternoonMeans16_1, aes(x=lat, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Early Afternoon Relative Humidity with Latitude") + 
  xlab("Latitude") + ylab("Relative Humidity (%)")

ggplot(AfternoonMeans16_2, aes(x=lat, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Late Afternoon Relative Humidity with Latitude") + 
  xlab("Latitude") + ylab("Relative Humidity (%)")

ggplot(AfternoonMeans17_1, aes(x=lat, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Early Afternoon Relative Humidity with Latitude") + 
  xlab("Latitude") + ylab("Relative Humidity (%)")

ggplot(AfternoonMeans17_2, aes(x=lat, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Late Afternoon Relative Humidity with Latitude") + 
  xlab("Latitude") + ylab("Relative Humidity (%)")

#rel humidity and long
ggplot(AfternoonMeans16_1, aes(x=long, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Early Afternoon Relative Humidity with Longitude") + 
  xlab("Longitude") + ylab("Relative Humidity (%)")

ggplot(AfternoonMeans16_2, aes(x=long, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Late Afternoon Relative Humidity with Longitude") + 
  xlab("Longitude") + ylab("Relative Humidity (%)")

ggplot(AfternoonMeans17_1, aes(x=long, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Early Afternoon Relative Humidity with Longitude") + 
  xlab("Longitude") + ylab("Relative Humidity (%)")

ggplot(AfternoonMeans17_2, aes(x=long, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Late Afternoon Relative Humidity with Longitude") + 
  xlab("Longitude") + ylab("Relative Humidity (%)")

#rel humidity and land use
ggplot(AfternoonMeans16_modLU_1, aes(x=as.factor(hour), y=RelHumid, fill=LandUse)) + 
  geom_boxplot() + ggtitle("2016 Early Afternoon Relative Humidity by Land Use") + 
  xlab("Hour") + ylab("Relative Humidity (%)")

ggplot(AfternoonMeans16_modLU_2, aes(x=as.factor(hour), y=RelHumid, fill=LandUse)) + 
  geom_boxplot() + ggtitle("2016 Late Afternoon Relative Humidity by Land Use") + 
  xlab("Hour") + ylab("Relative Humidity (%)")


ggplot(AfternoonMeans17_modLU_1, aes(x=as.factor(hour), y=RelHumid, fill=LandUse)) + 
  geom_boxplot() + ggtitle("2017 Early Afternoon Relative Humidity by Land Use") + 
  xlab("Hour") + ylab("Relative Humidity (%)")

ggplot(AfternoonMeans17_modLU_2, aes(x=as.factor(hour), y=RelHumid, fill=LandUse)) + 
  geom_boxplot() + ggtitle("2017 Late Afternoon Relative Humidity by Land Use") + 
  xlab("Hour") + ylab("Relative Humidity (%)")


#rel humidity and distance to ocean
ggplot(AfternoonMeans16_1, aes(x=Dist2Ocean, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Early Afternoon Relative Humidity with Distance to Ocean") + 
  xlab("Distance to Ocean (m)") + ylab("Relative Humidity (%)")

ggplot(AfternoonMeans16_2, aes(x=Dist2Ocean, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2016 Late Afternoon Relative Humidity with Distance to Ocean") + 
  xlab("Distance to Ocean (m)") + ylab("Relative Humidity (%)")

ggplot(AfternoonMeans17_1, aes(x=Dist2Ocean, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Early Afternoon Relative Humidity with Distance to Ocean") + 
  xlab("Distance to Ocean (m)") + ylab("Relative Humidity (%)")

ggplot(AfternoonMeans17_2, aes(x=Dist2Ocean, y=RelHumid, color=as.factor(hour))) + 
    geom_point() + ggtitle("2017 Late Afternoon Relative Humidity with Distance to Ocean") + 
  xlab("Distance to Ocean (m)") + ylab("Relative Humidity (%)")

```

##EDA - variogram

See if the spatial random effects approach is worth it

```{r variogram}
#Pick an hour of data to look at to make this a bit simpler
#thinking 17 since it is kind of between the shift in aft to evening temps
MidAftTemp16 <- HourlyMeans16[HourlyMeans16$hour == 17,]
MidAftTemp17 <- HourlyMeans17[HourlyMeans17$hour == 17,]

##start to look at varigrams for temp - 2016
varDatTemp16 <- as.geodata(MidAftTemp16, coords.col = 2:3, data.col=5)
plot(varDatTemp16)

v1Temp16 <- variog(varDatTemp16, trend="1st", option="bin")
plot(v1Temp16)
v1Temp16.c <- variog(varDatTemp16, op="cloud")
v1Temp16.bc <- variog(varDatTemp16, bin.cloud=TRUE)

plot(v1Temp16.c, main="variogram cloud 2016 Temp")
plot(v1Temp16.bc, bin.cloud=TRUE, main="clouds for binned variogram 2016 Temp")

##variogram for temp 2017
varDatTemp17 <- as.geodata(MidAftTemp17, coords.col = 2:3, data.col=5)
plot(varDatTemp17)

v1Temp17 <- variog(varDatTemp17, trend="1st", option="bin")
plot(v1Temp17)
v1Temp17.c <- variog(varDatTemp17, op="cloud")
v1Temp17.bc <- variog(varDatTemp17, bin.cloud=TRUE)

plot(v1Temp17.c, main="variogram cloud 2017 Temp")
plot(v1Temp17.bc, bin.cloud=TRUE, main="clouds for binned variogram 2017 Temp")



#variograms for RH - 2016
varDatRH16 <- as.geodata(MidAftRH16, coords.col = 2:3, data.col=6)
plot(varDatRH16)

v1RH16 <- variog(varDatRH16, trend="1st", option="bin")
plot(v1RH16)
v1RH16.c <- variog(varDatRH16, op="cloud")
v1RH16.bc <- variog(varDatRH16, bin.cloud=TRUE)

plot(v1RH16.c, main="variogram cloud 2016 Relative Humidity")
plot(v1RH16.bc, bin.cloud=TRUE, main="clouds for binned variogram 2016 Relative Humidity")



#varigrams for RH 2017
varDatRH17 <- as.geodata(MidAftRH17, coords.col = 2:3, data.col=6)
plot(varDatRH17)

v1RH17 <- variog(varDatRH17, trend="1st", option="bin")
plot(v1RH17)
v1RH17.c <- variog(varDatRH17, op="cloud")
v1RH17.bc <- variog(varDatRH17, bin.cloud=TRUE)

plot(v1RH17.c, main="variogram cloud 2017 Relative Humidity")
plot(v1RH17.bc, bin.cloud=TRUE, main="clouds for binned variogram 2017 Relative Humidity")

```

##Start in on regressions

So we would need to account for a few things
* spatial random effects
* fixed effects
  - land use
  - distance from ocean
  - hour
  
The relationship should change throughout the day, you see that the slopes of some of these relationships shift over time...
* hour x land use (really key)
* hour x distance2ocean? 

Alternatively, could run a few separate models at different times of day so you avoid some of the interactive effects
* early afternoon (12-14)
* peak temp (15-17)
* evening (18-21)

spBayes information
* https://faculty.ucr.edu/~jflegal/203/spBayes_tutorial.pdf
* https://cran.r-project.org/web/packages/spBayes/spBayes.pdf

If you do spDynLm I think you end up with different regressions for each hour
* so you end up with temps at each hour as your y variables
* then you coordinates are your spatial random effects
* your fixed effects are then land use and distance2ocean

```{r spatial temp}

#change your coordinates to UTM so the distances are consistent
coords_sf <- st_as_sf(coords, coords = c("X", "Y"), crs = 4326)
#need to check the UTM zone
coords_UTM <- st_transform(coords_sf, CRS("+proj=utm +zone=16 ellps=WGS84"))


##add knots to make this run faster
#https://rpubs.com/jimclark/751504

```